### 2014-04-14 template.tex
First round of editing

%%% template.tex
%%%
%%% This LaTeX source document can be used as the basis for your technical
%%% paper or abstract. Intentionally stripped of annotation, the parameters
%%% and commands should be adjusted for your particular paper - title, 
%%% author, article DOI, etc.
%%% The accompanying ``template.annotated.tex'' provides copious annotation
%%% for the commands and parameters found in the source document. (The code
%%% is identical in ``template.tex'' and ``template.annotated.tex.'')

\documentclass[conference]{acmsiggraph}

\TOGonlineid{45678}
\TOGvolume{0}
\TOGnumber{0}
\TOGarticleDOI{1111111.2222222}
\TOGprojectURL{}
\TOGvideoURL{}
\TOGdataURL{}
\TOGcodeURL{}

\title{A Triangle Mesh Approach to Painting}

\author{Mark D. Benjamin\thanks{e-mail:mdbenjam@princeton.edu} \hspace{10 pt} Princeton Unviersity\\ Adam Finkelstein\thanks{e-mail:af@cs.princeton.edu} \hspace{10 pt} Princeton University\\ Stephen Diverdi\thanks{e-mail:diverdi@google.com} \hspace{10 pt} Google}
\pdfauthor{Mark D. Benjamin, Adam Finkelstein, Stephen Diverdi}

\keywords{triangle mesh, painting, vector graphics}

\begin{document}

%% \teaser{
%%   \includegraphics[height=1.5in]{images/sampleteaser}
%%   \caption{Spring Training 2009, Peoria, AZ.}
%% }

\maketitle

\begin{abstract}

\end{abstract}

\begin{CRcatlist}
  %\CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
  %\CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
\end{CRcatlist}

\keywordlist

%% Use this only if you're preparing a technical paper to be published in the 
%% ACM 'Transactions on Graphics' journal.

\TOGlinkslist

%% Required for all content.

\copyrightspace

\section{Introduction}

All computer painting programs must store data through rasterization or vectorization. 
However, traditional computer painting programs take inputs of the same type as the stored data. The 
extremely popular Adobe programs Photoshop and Illustrator exemplify this. The former
allows a user to paint how he would on a physical medium. By letting the mouse act as
a brush, a user may color individual pixels on the screen. The latter lets a user paint
using geometric primitives such as lines and B\'{e}zier curves. The program then stores
these geometric primitives in order to render the drawing. Although using vector graphics
is likely less intuitive, it has some distinct advantages. First, there is no resolution
limit. One can zoom into an image as much as he would like and still see crisp edges.
Second, certain editing operations may be easier to perform on vector graphics.

In this paper we introduce TrianglePainter which  
uses a slightly different representation than traditional raster or vector graphics.
We instead use triangle meshes to store the data of the painting.
Since triangles are a geometric primitive used in vector graphics our approach achieves the
same advantages traditional vector graphics have. However, our representation allows the
user to paint with vector graphics without worrying about the underlying implementation.

In order to paint in our program the user simply drags his mouse on the screen to make
strokes akin to the process in a raster graphics program. Upon mouse-up the stroke is
converted to our underlying triangle mesh representation. In this way a user may paint using
vector graphics without worrying about their representation.

Furthermore, our representation allows painting at any scale. A user can make large strokes
while zoomed out and then zoom in to make fine details. All of this can happen without
loss of quality since the data is stored using triangles, not in pixels.

To do the transformation we use a combination of rasterization and marching squares to
determine the contour of the stroke. From the contour we can create a Delaunay triangulation
to represent the stroke. By then introducing a merging process we can take the triangulation
of the new stroke and combine it with the triangulation of the existing canvas. This yields
a new canvas with all the strokes combined.

TrianglePainter reduces the barriers in creating vector graphics. This in turn allows 
artists to focus on the painting and not the way in which they paint, while still 
getting the editing benefits of vector graphics.

\section{Triangulating Strokes}

Here we consider how to convert a user's cursor motion into a triangle-mesh representation
of their stroke.

\subsection{Converting Strokes to Contours}
The contour(s) define the outline of the stroke. Simply connected strokes, those without holes, have only 
one outer contour, whereas strokes with holes, such as a circle or 
figure-eight, will have more than one.
Converting a user's mouse motion into a contour requires two steps.
First, the user's mouse motion is captured by a set of polygons which can be
rendered to yield a rasterized version of the stroke.
Second, to extract the contours the stroke is rasterized in black and marching squares
is run to extract iso-contours of 50\% gray. This returns a very dense set of points 
which we reduce through a pruning process to
get a sparse but accurate representation of the contour.

\\

What do you think about this?

Note that despite the reliance on rasterization it is still possible by using
anti-aliasing and certain brushes to get sub-pixel resolution.

\subsection{Converting Contours to Triangles}
The contour contains two important pieces of information. First, it describes all of the points of the stroke.
Second, it describes the boundaries of the stroke. A triangle mesh must contain all the points and stay
within the boundaries. For example, a concave contour should not contain triangles inside the concavity.
Furthermore, these boundaries will be crucial when compositing a stroke onto a canvas.

We utilized Jonathan Shewchuk's Triangle library to perform Delaunay triangulations
on the given set of points and boundaries. The library outputs a triangle mesh that
approximately represents the drawn stroke.

\subsection{Feathered Strokes}
Finally, we must note a few changes to the above procedure in the case of feathered strokes.
Feathered strokes have two pieces, a normal inner stroke, and a ring around this stroke
where the color varies from that defined on the inner stroke to completely transparent. We
represent these strokes with two sets of points and boundaries. One set which defines the
normal inner stroke, and one set that defines the outside transparent edge of the stroke.

In order to obtain this information we must change the way we capture strokes. We accomplish
this by drawing two sets of parallelograms. One for the inner stroke and one for the outer.
When we rasterize the strokes, we do the outer one first with a color of 50\% gray and
then do the inner one second with a color of black. We run marching squares to find the
iso-contours of 25\% gray and 75\% gray and then use these two iso-contours to inform the
two sets of points and edges representing the stroke. At this point we can triangulate
the stroke normally. By providing the constraint edges and points the Triangle library will
ensure that the triangles of the outer and inner strokes do not cross the boundaries specified.

\section{Compositing Strokes}
Once the stroke has been converted to triangles by the above procedure it is necessary to
composite the stroke onto the canvas. Before drawing, every canvas is triangulated
into two large triangles. Compositing a stroke requires combining the points and boundaries
from the old strokes already on the canvas and the points and boundaries from the new stroke.
Before we get to the actual compositing, we need to understand how the triangles are colored.

\subsection{How Colors are Stored}
As we will subsequently see, the triangles on the canvas are not persistent, but the
points and boundaries from each stroke are. Therefore, all color information must be
stored in the points. All of the boundaries specify a color boundary as well. When a
stroke is drawn, on the inside of the boundary we expect the triangles to take on the
color of the stroke. On the outside of the boundary we expect the triangles to take
on the color of whatever previous strokes were laid down. In order to keep color information
stored in the points that describes the colors of the triangles, we introduce the
notion of color arcs.

A point's color arc is a color associated with a set of two vectors such that the point 
will take on the color for any triangle it is a part of whose centroid lies within the 
color arc. Points can have multiple color arcs. The union of all of the color arcs should
cover all $2\pi$ radians around the point. And no two color arcs should intersect.
In this way for any triangle formed with a given point, the point's color assignment within
that triangle can be uniquely determined by finding which color arc the centroid is in.

\subsection{Finding Modified Points}
One way to create our new triangulation of the canvas is to add our points and boundaries
to the old list of points and boundaries and have Triangle perform a Delaunay triangulation
of the entire set. Since this triangulation will preserve the boundaries we provided,
(i.e. no triangles will cross any of the constraint edges) all of the strokes will be
defined by a set of triangles. There is clearly a way to color these triangles to
correctly represent the strokes. We discuss how to find the colors in subsequent sections.

However, a canvas may easily contain thousands of triangles, and a new stroke may only 
intersect some small subset of them. It seems wasteful in terms of computational
resources to perform a re-triangulation of the entire canvas when most of the triangles
are not affected by this new stroke. This also means we can't keep the effects of
strokes local. In other words a stroke in the bottom left may have an effect in the top
right.

In order to solve this problem our program selectively choses which triangles can remain
and which triangles need to be included in the re-triangulation. We make a 20-by-20 grid
and mark all of the cells which contain one a triangle from the new stroke. We then
find the set of triangles from the old triangulation that intersect these grid squares.
The set of edges from these triangles that do not intersect these grid cells form a ring
around the area of our new triangles. We make all of the edges in this ring into constraints
in our triangulation. All of the points and constraint edges from the old triangulation are
included as well as the points and constraint edges from the new triangulation. Triangulating
these points and edges gives the new geometry for the area we're concerned with, but due
to the ringed constraint edges it has no effect on any other part of the canvas. Therefore
we can reuse the old triangles from the rest of the canvas. Together the new triangulation
plus the old triangle gives a new triangle mesh representation of the entire canvas.

\subsection{Determining Colors}
The final piece of the algorithm is to determine the color arcs for each of the points.
When first putting down points from a stroke, color arcs are immediately assigned. For
a normal stroke there are two color arcs. The first arc has the color of the current
brush and covers the inside of the stroke. The second arc has the color of whatever color was
already there before this point and covers the outside of the stroke. To determine
what color was already there we do a point in triangle test on the current canvas. This
gives us the triangle, and since we know the color arcs of the three vertices, we can calculate
with bilinear interpolation the color at the concerned point.

For a feathered stroke points only have one color arc. For the inner stroke the color arcs
go all $2\pi$ radians around and have the color of the stroke. For the outer stroke the
color arcs go all $2\pi$ radians around and are whatever color was there before. This means
that all triangles between the inner and outer strokes will have a linear change in color
from this stroke's color, to that of the surrounding.

Whenever a new stroke is composited on an old stroke there will be a change in the color of
any point that is now covered. The vectors defining the arc remain the same, but the color
itself will be the new color of the stroke over the old color of the arc. In the case of a
feathered stroke, where some parts of the stroke have changing color, the color of that
triangle at that point must be calculated by bilinear interpolation.

Finally, the most interesting case is when constraint edges intersect. When performing the 
triangulation of the new stroke on top of the old strokes, in most situations constraint
edges will intersect. Since no triangle can pass through these edges it is necessary for a point
to be added at the intersection. The Triangle library does this automatically. However, this
intersection point has no color data since we did not create it when we created the
stroke.

To determine the color data for an intersection point it is necessary to follow the constraint
edges to find points with color data. Note that the neighboring points may be intersection
points as well and therefore not have any color data. Once we've found colored points we
linearly interpolate the color arcs from either side of our intersection point to determine 
the colors at our new point. Once we have the average color arcs from both directions we
composite the color arcs of the new stroke over that of the old stroke in order to get four
color arcs of the four combinations of the old color arcs.

\subsection{Discussion}
Our work with this algorithm has lead to observations in three main areas.

\subsubsection{Stability}
Something we've worked hard to achieve is program stability. Certain geometries are fatal
for the Triangle library. The current instability in the program is normally caused by
providing the library with such difficult geometries. To improve program stability would
require addressing geometries such as vertices that are too close together. This would
take more processing time and is something we have yet to explore. However, clearly this
problem must be solved before the program could be used formally.

\subsubsection{Performance}
The current implementation uses Python and PyOpenGL. Python's performance is worse than
a lower level implementation in C++ would be. The drawing of the canvas is quick with
no lag. However, the actual process to extract a stroke from rasterization to triangle
representation on the canvas is slow. On a blank canvas strokes take on the order of
1 second to process. However, on a canvas with lots of geometry a stroke that covers
that geometry can take several seconds to render. Some of this time is due to poor
implementation of the algorithm. However, there are lots of calculations
to do with the old points affected by the new stroke and the points associated with the
new stroke itself. Since these geometries can get arbitrarily complicated, these
calculations can take an arbitrarily long. This brings us to the final point, scalability.

\subsubsection{Scalability}
No matter how good the implementation, the fact that paintings can get arbitrarily complicated
means there must be some way to simplify the mesh. Although we have yet to implement
such a simplification algorithm, we have discussed the possibility. In most situations where
a lot of strokes are composited on top of one another, most of the geometry is redundant. 
For example, when an opaque stroke is laid down on top of a complex geometry, that complex
geometry no longer has any useful information since the outline of the opaque stroke
completely defines it. Furthermore, even with translucent or feathered strokes, enough
composited strokes may produce complex geometry that does not add much to the output.
Some of this geometry can be reduced without noticeable changes in the output. Such a
simplification algorithm must be implemented to ensure a painter can continue to paint
without the program slowing to an unusable state.

\section{Results}

\section{Conclusion}

\section*{Acknowledgements}

\bibliographystyle{acmsiggraph}
\bibliography{template}
\end{document}

### 2014-04-19 template.tex
Before writing discussion and results

@@
 \documentclass[conference]{acmsiggraph}
-
+\usepackage{subcaption}
 \TOGonlineid{45678}
@@
 converted to our underlying triangle mesh representation. In this way a user may paint using
-vector graphics without worrying about their representation.
+vector graphics without worrying about the representation.
 
@@
 
+\section{Related Works}
+
+Previous work has focused on converting images into vector graphics to take advantage of the
+efficient storage, easy editing, and infinite resolution it provides. For example \cite{lecot:ARD:2006}
+minimizes an energy function to segment an image into a number of regions that are bounded
+by cubic splines and filled with solid colors or gradients. A different approach by \cite{Lai:2009:ATG:1531326.1531391} 
+automatically creates gradient meshes with support for holes. Yet another example is \cite{10.1109/TVCG.2012.76} 
+which converts an image into a triangle mesh. All three of these can convert images into
+a vector representation, yet none of them allow a user to create a painting in the vector medium.
+
+Other work has proposed novel vector graphic data structures. For example \cite{Frisken:2000:ASD:344779.344899}
+defines Adaptively Sampled Distance Fields (ADF). ADFs specify a signed distance function to a surface.
+to render the shape requires sampling the function to determine whether or not a pixel is on the surface.
+A quadtree accompanies the distance function to specify where higher sampling rates are necessary. \cite{Bremer:2001:VCM}
+further describes how ADFs can be created and used. However, it is unclear how ADFs might be used in a general
+painting program where strokes with solid colors and gradients are composited on top of one another.
+
+Another series of works have created entirely new ways of painting. Diffusion curves as specified in \cite{Orzan:2008:DCV:1360612.1360691}
+describe a new painting technique where the artists specify edges and colors for those edges. Their system
+then solves a Poisson equation to diffuse the colors between boundary conditions specified at the edges.
+The algorithm produces some truly beautiful drawing. Similarly \cite{McCann:2008:RGP:1360612.1360692}
+lets artists paint in the gradient domain. Both of these approaches yield fantastic works of art, yet
+they are really new mediums for artists to explore rather than a vectorized version of conventional
+painting techniques.
+
+Work has also been done in multi-resolution painting using raster graphics. \cite{Berman:1994:MPC:192161.192181}
+stores image information in a quadtree so that different parts of the image can have different levels of detail.
+This means the image effectively has an infinite level of detail, yet since it is stored in raster form
+there are still limitations. For example a stroke painted at coarse detail will still look blurry when
+zoomed in. The strokes are limited to the resolution they are painted at. \cite{Carr:2004:PD:1186562.1015809}
+also allows the development of multi-resolution images using rasters, and as such suffers from the
+same limitations. Another multi-resolution approach by \cite{Perlin:1995:LPP:218380.218437} uses
+procedural textures to avoid any fundamental limit on resolution. Although useful for textures,
+generating procedural textures is difficult and does not help artists easily create.
+
+Our work builds on these findings. However, our goal is to enable artists to create vector
+graphics through standard painting techniques without having to deal with the underlying representation.
+We use the triangle mesh structure described in \cite{10.1109/TVCG.2012.76} to build a painting program
+where the triangle mesh is created while the artists lays down strokes.
+
+\section{The Canvas Model}
+The data necessary to render the painting resides in the canvas. However, the canvas contains more than just
+a triangle mesh and understanding the underlying data structures is important for 
+understanding the rest of the algorithm.
+
+\subsection{Triangle Mesh}
+When a user paints he generates points that are stored on the canvas. These points are
+connected to form a triangle mesh. In every triangle the vertices are colored to either
+produce a solid color or a linear gradient. In fact a given point that is part of
+multiple triangles can take on different colors in each triangle. Note that although
+the points are persistent, the triangles are not.
+
+\subsection{Boundary Edges}
+Although the triangles may change over time, they must abide by constraints known as
+boundary edges. A boundary edge connects two points and specifies that no triangle
+can cross it. Like points, boundary edges are persistent. So as triangles come and go they
+may never cross a boundary edges.
+
+\subsection{Results}
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.4\textwidth]{images/stroke}
+    \caption{An example triangulation. Blue lines are boundary edges. Black lines form the triangle
+    mesh with the blue lines.}
+\end{figure}
+
+\subsection{Color Arcs}
+Since triangles are not persistent they cannot store color data.
+Storing colors in points works well when the color is the same for all triangles connected
+to that point. However, when a point lies on a sharp boundary it may take on different colors in
+triangles on either side of the boundary. The need for the same point to take on different
+colors in different triangles lead to the development of color arcs.
+
+A color arc has three components: a start vector, an end vector, and a color. 
+Determining the color of a vertex in a given triangle requires two steps. First,
+the vector from the point to the centroid of the triangle is calculated. Second,
+the color arc is found which contains this vector. Containing the vector means one must turn
+clockwise from the starting vector to reach it and counter-clockwise from the ending vector.
+The point takes on the color of the arc in which the vector from the point to the centroid resides.
+
+The color arcs for every point must satisfy three conditions. First, they must be disjoint. Second,
+they must cover all $2\pi$ radians around the point. These first two conditions ensure that for any
+triangle a point will take on exactly one color. Third, the start and end vectors must line up
+with boundary edges. This condition makes all colors and gradients emanate from boundaries.
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.4\textwidth]{images/colorarcs}
+    \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
+    mesh with the boundary edges. Each point is given a color arc which defines its color in that range.
+    Note that the points in the middle row have different colors on top and bottom. This gives blue
+    triangles on top and red on bottom with a hard boundary in between.}
+\end{figure}
+
+\subsection{Rendering}
+Once the triangle mesh has been generated it is simple to render. In each triangle the vertices
+take on a given color based on their color arcs and the centroid of the triangle. If the three
+vertices in the triangle take on the same color value then the triangle will be solid. If instead the 
+vertices take on different color values then there will be a gradient over the triangle.
+
+Since the conditions on color arcs make colors emanate from boundaries it's easy to create regions
+containing either a solid color or linear gradient. A region will have a solid color if the
+boundaries that surround it all take on the same color. If instead a region is surrounded by
+two boundaries, one which has a given color and the other which has a transparent version of that color
+then there will be a linear gradient in that region between the two boundaries.
+
+\includegraphics[height=1.5in]{images/softstrokeboundaryedges}
+
 \section{Triangulating Strokes}
@@
 \subsection{Converting Strokes to Contours}
-The contour(s) define the outline of the stroke. Simply connected strokes, those without holes, have only 
-one outer contour, whereas strokes with holes, such as a circle or 
-figure-eight, will have more than one.
-Converting a user's mouse motion into a contour requires two steps.
-First, the user's mouse motion is captured by a set of polygons which can be
+A set of contours define the outline of a stroke. For simply connected strokes, those without holes, 
+one contour describes the whole stroke. However, strokes with holes, such as a circle or 
+figure-eight, will have more than one contour.
+Converting a user's mouse motion into a set of contours requires two steps.
+First, the user's mouse movements are captured by a set of polygons which can be
 rendered to yield a rasterized version of the stroke.
-Second, to extract the contours the stroke is rasterized in black and marching squares
+Second, to obtain the contours the stroke is rasterized in black and marching squares
 is run to extract iso-contours of 50\% gray. This returns a very dense set of points 
@@
 
-\\
+Adam and Steve should we talk about how anti-aliasing + marching squares can achieve sub-pixel
+resolution?
 
-What do you think about this?
-
-Note that despite the reliance on rasterization it is still possible by using
-anti-aliasing and certain brushes to get sub-pixel resolution.
+% Note that despite the reliance on rasterization it is still possible by using
+% anti-aliasing and certain brushes to get sub-pixel resolution.
 
@@
 The contour contains two important pieces of information. First, it describes all of the points of the stroke.
-Second, it describes the boundaries of the stroke. A triangle mesh must contain all the points and stay
-within the boundaries. For example, a concave contour should not contain triangles inside the concavity.
-Furthermore, these boundaries will be crucial when compositing a stroke onto a canvas.
+Second, segments connecting adjacent points describe the boundaries of the stroke. 
+A triangle mesh must contain all the points and stay
+within the boundaries. For example, a concave contour should not contain triangles inside the concavity since such triangles
+would be outside of the boundary.
+Furthermore, these boundaries will later be crucial when compositing a stroke onto a canvas.
 
 We utilized Jonathan Shewchuk's Triangle library to perform Delaunay triangulations
-on the given set of points and boundaries. The library outputs a triangle mesh that
+on the given set of points and boundaries. The library outputs a mesh that
 approximately represents the drawn stroke.
 
-\subsection{Feathered Strokes}
-Finally, we must note a few changes to the above procedure in the case of feathered strokes.
-Feathered strokes have two pieces, a normal inner stroke, and a ring around this stroke
-where the color varies from that defined on the inner stroke to completely transparent. We
-represent these strokes with two sets of points and boundaries. One set which defines the
-normal inner stroke, and one set that defines the outside transparent edge of the stroke.
+\begin{figure*}
+\centering
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardrendered}
+  \caption{A subfigure}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardpruned}
+  \caption{A subfigure}
+  \label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardmesh}
+  \caption{A subfigure}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{A figure with two subfigures}
+\label{fig:test}
+\end{figure*}
 
-In order to obtain this information we must change the way we capture strokes. We accomplish
-this by drawing two sets of parallelograms. One for the inner stroke and one for the outer.
-When we rasterize the strokes, we do the outer one first with a color of 50\% gray and
-then do the inner one second with a color of black. We run marching squares to find the
-iso-contours of 25\% gray and 75\% gray and then use these two iso-contours to inform the
-two sets of points and edges representing the stroke. At this point we can triangulate
-the stroke normally. By providing the constraint edges and points the Triangle library will
-ensure that the triangles of the outer and inner strokes do not cross the boundaries specified.
+\begin{figure*}
+\centering
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softrendered}
+  \caption{A subfigure}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softpruned}
+  \caption{A subfigure}
+  \label{fig:sub2}
+\end{subfigure}
+\begin{subfigure}[b]{0.3\textwidth}
+  \centering
+  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softmesh}
+  \caption{A subfigure}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{A figure with two subfigures}
+\label{fig:test}
+\end{figure*}
+
+% \includegraphics[height=1.5in]{images/triangulationprocess}
+
+\subsection{Soft Strokes}
+The above procedure converts a hard brush into a triangle mesh. Yet, a somewhat more interesting
+case is converting a soft brush with linear gradient into a triangle mesh. A soft stroke has two
+components, an inner hard stroke and an outer gradient stroke. Since the triangles in the inner and
+outer parts of the stroke must be colored differently, none of the triangles may cross the boundary
+between the hard and soft regions.
+
+This representation needs a slightly different procedure. The user's mouse motion
+is capture in two sets of polygons. The first set describes the outer soft stroke and the second set describes
+the inner hard stroke. Next the outer stroke polygons are rendered in 50\% gray and the inner stroke
+is rendered in black. This means the boundary of the outer stroke is the 25\% gray iso-contour and the
+boundary of the inner stroke is the 75\% gray iso-contour.
+
+Again, using marching squares these contours are extracted and then pruned. The boundaries and points
+derived from the contours are fed into Triangle and a mesh is returned. This mesh represents the soft
+stroke where no triangle crosses the boundary between the soft and hard parts
+of the stroke.
+
+\includegraphics[height=1.5in]{images/softstroke}
+\includegraphics[height=1.5in]{images/hardvssoft}
 
@@
 Once the stroke has been converted to triangles by the above procedure it is necessary to
-composite the stroke onto the canvas. Before drawing, every canvas is triangulated
-into two large triangles. Compositing a stroke requires combining the points and boundaries
-from the old strokes already on the canvas and the points and boundaries from the new stroke.
-Before we get to the actual compositing, we need to understand how the triangles are colored.
+composite the stroke onto the canvas. We first consider the canvas model and then how a
+stroke can be composited onto the canvas.
 
-\subsection{How Colors are Stored}
-As we will subsequently see, the triangles on the canvas are not persistent, but the
-points and boundaries from each stroke are. Therefore, all color information must be
-stored in the points. All of the boundaries specify a color boundary as well. When a
-stroke is drawn, on the inside of the boundary we expect the triangles to take on the
-color of the stroke. On the outside of the boundary we expect the triangles to take
-on the color of whatever previous strokes were laid down. In order to keep color information
-stored in the points that describes the colors of the triangles, we introduce the
-notion of color arcs.
+\subsection{Adding Points to the Canvas}
 
-A point's color arc is a color associated with a set of two vectors such that the point 
-will take on the color for any triangle it is a part of whose centroid lies within the 
-color arc. Points can have multiple color arcs. The union of all of the color arcs should
-cover all $2\pi$ radians around the point. And no two color arcs should intersect.
-In this way for any triangle formed with a given point, the point's color assignment within
-that triangle can be uniquely determined by finding which color arc the centroid is in.
+Every canvas begins with four points, one at each corner of the screen, and four boundaries, one
+for each edge of the screen. These points are originally part of two big triangles which
+divide the canvas.
+
+In the simplest approach the points and boundaries from the new stroke are added
+to the list of points and boundaries already stored. All of this data can be fed into Triangle
+to produce a new mesh. The new mesh contains all of the points of the old and new strokes. Since
+all of the boundaries are stored and fed into Triangle the stroke edges are preserved. This means that
+no triangle can cross the boundary of a stroke. Without this constraint it would be impossible to color the
+triangles since they would straddle color arcs. Once a new mesh is formed the color arcs of the 
+new and old points must be computed.
+
+\includegraphics[height=1.5in]{images/canvastriangulation}
+
+\subsection{Determining Colors}
+The color arcs for new points depend on two values, the color of the stroke and the color
+of the canvas at that point. The points on a hard stroke will have two color arcs. One
+arc describes the inside of the stroke. The color for this arc is the color of the stroke
+composited over the color of the canvas at that point. The other arc describes the outside
+of the stroke. The color for this arc is just the color of the canvas at that point. The
+points on a soft stroke will have one color arc. The color arc for points on the inner
+contour is the color of the stroke composited over the color of the canvas at that point.
+The color arc for points on the outer contour is the color of canvas at that point.
+This gives any triangle connecting inside and outside contour points a linear gradient
+from the stroke color composited over the canvas color to the canvas color.
+
+The color arcs for old points remain the same unless part of the new stroke covers it.
+In this case the colors in each arc must be replaced with the stroke color composited over
+the old color of the arc. If the new stroke is soft then the color at the location of
+the old point must be determined through bilinear interpolation before compositing.
+
+\subsection{Intersection Points}
+
+When Triangle is asked to triangulate a set of points with overlapping boundaries it must
+insert a point at the intersection, since without it at least one triangle would have to
+cross a boundary. These points are introduced with no color information, but the final colors for 
+all of the other points in the mesh have already been determined.
+
+An intersection point lies on two boundaries and therefore has four boundary edges emanating from it. 
+To determine the colors of an intersection points requires knowing the colors associated with
+points on the boundaries. For both boundaries travel both directions to the first point with color data. Note this may not
+be the first point on each side since this intersection point may be adjacent to other intersection
+points without color data. For both boundaries linearly interpolate the colors in the color arcs. 
+This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
+The colors from the new stroke are composited over the colors from the old stroke to yield four new
+colors, one for each region defined by the boundary intersections.
+
+\includegraphics[height=1.5in]{images/determiningcolors}
 
 \subsection{Finding Modified Points}
-One way to create our new triangulation of the canvas is to add our points and boundaries
-to the old list of points and boundaries and have Triangle perform a Delaunay triangulation
-of the entire set. Since this triangulation will preserve the boundaries we provided,
-(i.e. no triangles will cross any of the constraint edges) all of the strokes will be
-defined by a set of triangles. There is clearly a way to color these triangles to
-correctly represent the strokes. We discuss how to find the colors in subsequent sections.
+The above procedure has two important drawbacks. First, it processes many points that don't need
+consideration which takes a considerable amount of processing time. Second, since it acts on the whole
+canvas, local changes can have global effects. To avoid this problem the algorithm determines
+which triangles can remain
+and which triangles need to be included in the re-triangulation. To accomplish this a static grid
+is generated and all of the cells which contain a triangle from the new stroke are marked. 
+The set of triangles from the old triangulation that intersect these grid squares are found.
+The set of edges from these triangles that do not intersect these grid cells form a ring
+around the area of our new triangles. The edges in this ring become constraints
+in the triangulation. All of the points and constraint edges from the old triangulation are
+included as well as the points and constraint edges from the new triangulation. Triangulating
+these points and edges gives the new geometry for the modified area and has no effect on any
+other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused. 
+The new triangulation combined with the old triangles give a new triangle mesh representation of the entire canvas.
 
-However, a canvas may easily contain thousands of triangles, and a new stroke may only 
-intersect some small subset of them. It seems wasteful in terms of computational
-resources to perform a re-triangulation of the entire canvas when most of the triangles
-are not affected by this new stroke. This also means we can't keep the effects of
-strokes local. In other words a stroke in the bottom left may have an effect in the top
-right.
+\includegraphics[height=1.5in]{images/findingmodifiedpoints}
 
-In order to solve this problem our program selectively choses which triangles can remain
-and which triangles need to be included in the re-triangulation. We make a 20-by-20 grid
-and mark all of the cells which contain one a triangle from the new stroke. We then
-find the set of triangles from the old triangulation that intersect these grid squares.
-The set of edges from these triangles that do not intersect these grid cells form a ring
-around the area of our new triangles. We make all of the edges in this ring into constraints
-in our triangulation. All of the points and constraint edges from the old triangulation are
-included as well as the points and constraint edges from the new triangulation. Triangulating
-these points and edges gives the new geometry for the area we're concerned with, but due
-to the ringed constraint edges it has no effect on any other part of the canvas. Therefore
-we can reuse the old triangles from the rest of the canvas. Together the new triangulation
-plus the old triangle gives a new triangle mesh representation of the entire canvas.
-
-\subsection{Determining Colors}
-The final piece of the algorithm is to determine the color arcs for each of the points.
-When first putting down points from a stroke, color arcs are immediately assigned. For
-a normal stroke there are two color arcs. The first arc has the color of the current
-brush and covers the inside of the stroke. The second arc has the color of whatever color was
-already there before this point and covers the outside of the stroke. To determine
-what color was already there we do a point in triangle test on the current canvas. This
-gives us the triangle, and since we know the color arcs of the three vertices, we can calculate
-with bilinear interpolation the color at the concerned point.
-
-For a feathered stroke points only have one color arc. For the inner stroke the color arcs
-go all $2\pi$ radians around and have the color of the stroke. For the outer stroke the
-color arcs go all $2\pi$ radians around and are whatever color was there before. This means
-that all triangles between the inner and outer strokes will have a linear change in color
-from this stroke's color, to that of the surrounding.
-
-Whenever a new stroke is composited on an old stroke there will be a change in the color of
-any point that is now covered. The vectors defining the arc remain the same, but the color
-itself will be the new color of the stroke over the old color of the arc. In the case of a
-feathered stroke, where some parts of the stroke have changing color, the color of that
-triangle at that point must be calculated by bilinear interpolation.
-
-Finally, the most interesting case is when constraint edges intersect. When performing the 
-triangulation of the new stroke on top of the old strokes, in most situations constraint
-edges will intersect. Since no triangle can pass through these edges it is necessary for a point
-to be added at the intersection. The Triangle library does this automatically. However, this
-intersection point has no color data since we did not create it when we created the
-stroke.
-
-To determine the color data for an intersection point it is necessary to follow the constraint
-edges to find points with color data. Note that the neighboring points may be intersection
-points as well and therefore not have any color data. Once we've found colored points we
-linearly interpolate the color arcs from either side of our intersection point to determine 
-the colors at our new point. Once we have the average color arcs from both directions we
-composite the color arcs of the new stroke over that of the old stroke in order to get four
-color arcs of the four combinations of the old color arcs.
-
-\subsection{Discussion}
+\section{Discussion and Results}
 Our work with this algorithm has lead to observations in three main areas.
 
-\subsubsection{Stability}
+\subsection{Stability}
 Something we've worked hard to achieve is program stability. Certain geometries are fatal
@@
 
-\subsubsection{Performance}
-The current implementation uses Python and PyOpenGL. Python's performance is worse than
+\subsection{Performance}
+The current implementation uses Python and PyOpenGL. Python's performance is slower than
 a lower level implementation in C++ would be. The drawing of the canvas is quick with
@@
 
-\subsubsection{Scalability}
+\subsection{Scalability}
 No matter how good the implementation, the fact that paintings can get arbitrarily complicated
 means there must be some way to simplify the mesh. Although we have yet to implement
-such a simplification algorithm, we have discussed the possibility. In most situations where
+such a simplification algorithm, we have a couple suggestions on how it might be done. In most situations where
 a lot of strokes are composited on top of one another, most of the geometry is redundant. 
@@
 
-\section{Results}
+\subsection{Complexity}
+We now examine the growth of triangles, points, and boundaries on the canvas.
+
+% \begin{tabular}{| l | l | l | l|}
+%   \hline                       
+%   Number of Strokes & Triangles & Points & Boundaries \\
+%   \hline                       
+%   1 & 360 & 229 & 229 \\
+%   2 & 506 & 302 & 310 \\
+%   3 & 652 & 375 & 403 \\
+%   4 & 798 & 448 & 488 \\
+%   5 & 972 & 535 & 603 \\
+%   6 & 1164 & 631 & 713 \\
+%   \hline  
+% \end{tabular}
+% \includegraphics[height=1.5in]{images/tri360points229edges229}
+% \includegraphics[height=1.5in]{images/tri1164points631edges713}
+
+The complexity of the triangles increase slightly more than linearly since extra points and 
+boundaries are made when boundaries intersect.
+
+NOTES:
+
+-Maybe do the same analysis with soft strokes
+
+-Look at the time it takes to render these strokes
+
+-Then show more robust drawings and the number of triangles, points, and boundaries
 

### 2014-04-20 template.tex
Draft for edits

@@
 \begin{abstract}
+Vector graphics have a compact representation, are easy to manipulate, and have infinite resolution.
+Traditionally, vector graphics have required artists to master a new skill in order to effectively utilize
+them. Recent work has converted raster graphics to vector graphics, allowed multi-resolution raster painting,
+and created new ways to represent vector graphics. However, none of these have solved the problem of
+letting an artist use standard and intuitive raster painting techniques to generate vector graphics.
+Our algorithm, TrianglePainter, converts a user's strokes into a triangle mesh vector representation. 
+TrianglePainter lets artists focus on the art, while it handles the vectorization in the background.
 
@@
 All computer painting programs must store data through rasterization or vectorization. 
-However, traditional computer painting programs take inputs of the same type as the stored data. The 
-extremely popular Adobe programs Photoshop and Illustrator exemplify this. The former
-allows a user to paint how he would on a physical medium. By letting the mouse act as
-a brush, a user may color individual pixels on the screen. The latter lets a user paint
+However, traditional computer painting programs take inputs of the same type as the stored data. 
+For example, programs like Photoshop and GIMP
+allow a user to paint how he would on a physical medium. By letting the mouse act as
+a brush, a user may color individual pixels on the screen. By contrast, programs like 
+Illustrator and Inkscape let a user paint
 using geometric primitives such as lines and B\'{e}zier curves. The program then stores
 these geometric primitives in order to render the drawing. Although using vector graphics
-is likely less intuitive, it has some distinct advantages. First, there is no resolution
-limit. One can zoom into an image as much as he would like and still see crisp edges.
-Second, certain editing operations may be easier to perform on vector graphics.
+is likely less intuitive, it has some distinct advantages including a compact
+representation, infinite resolution, and easier manipulation.
 
@@
 uses a slightly different representation than traditional raster or vector graphics.
-We instead use triangle meshes to store the data of the painting.
-Since triangles are a geometric primitive used in vector graphics our approach achieves the
-same advantages traditional vector graphics have. However, our representation allows the
-user to paint with vector graphics without worrying about the underlying implementation.
+TrianglePainter uses a triangle mesh to store the data of the painting.
+Since triangles are a geometric primitive used in vector graphics this approach achieves the
+same advantages traditional vector graphics have. However, the representation allows the
+user to paint with vector graphics without dealing directly with the underlying implementation.
 
@@
 
-Furthermore, our representation allows painting at any scale. A user can make large strokes
+Furthermore, this representation allows painting at any scale. A user can make large strokes
 while zoomed out and then zoom in to make fine details. All of this can happen without
@@
 
-To do the transformation we use a combination of rasterization and marching squares to
-determine the contour of the stroke. From the contour we can create a Delaunay triangulation
-to represent the stroke. By then introducing a merging process we can take the triangulation
-of the new stroke and combine it with the triangulation of the existing canvas. This yields
-a new canvas with all the strokes combined.
+To transform a user's mouse inputs to a triangle mesh
+TrianglePainter uses a combination of rasterization and marching squares to
+find stroke contours. From the contour it performs a Delaunay triangulation
+and then merges this triangulation
+of the new stroke with the triangle mesh of the existing canvas. This process yields
+a new canvas with the combined strokes.
 
-TrianglePainter reduces the barriers in creating vector graphics. This in turn allows 
+TrianglePainter reduces the barrier to creating vector graphics. This in turn allows 
 artists to focus on the painting and not the way in which they paint, while still 
-getting the editing benefits of vector graphics.
+getting the benefits of vector graphics.
 
@@
 defines Adaptively Sampled Distance Fields (ADF). ADFs specify a signed distance function to a surface.
-to render the shape requires sampling the function to determine whether or not a pixel is on the surface.
+Rendering the shape requires sampling the function to determine whether or not a pixel is on the surface.
 A quadtree accompanies the distance function to specify where higher sampling rates are necessary. \cite{Bremer:2001:VCM}
@@
 then solves a Poisson equation to diffuse the colors between boundary conditions specified at the edges.
-The algorithm produces some truly beautiful drawing. Similarly \cite{McCann:2008:RGP:1360612.1360692}
+Similarly \cite{McCann:2008:RGP:1360612.1360692}
 lets artists paint in the gradient domain. Both of these approaches yield fantastic works of art, yet
@@
 same limitations. Another multi-resolution approach by \cite{Perlin:1995:LPP:218380.218437} uses
-procedural textures to avoid any fundamental limit on resolution. Although useful for textures,
-generating procedural textures is difficult and does not help artists easily create.
+procedural textures to avoid any fundamental limit on resolution. Although useful,
+generating procedural textures is difficult and does not help most artists easily create.
 
@@
 The data necessary to render the painting resides in the canvas. However, the canvas contains more than just
-a triangle mesh and understanding the underlying data structures is important for 
+a triangle mesh, and understanding the underlying data structures is important for 
 understanding the rest of the algorithm.
@@
 boundary edges. A boundary edge connects two points and specifies that no triangle
-can cross it. Like points, boundary edges are persistent. So as triangles come and go they
+can cross it. Like points, boundary edges are persistent. So as new triangles are created they
 may never cross a boundary edges.
-
-\subsection{Results}
 
@@
 
+\begin{figure}
+    \centering
         \includegraphics[height=1.5in]{images/softstrokeboundaryedges}
+    \caption{Place Holder.}
+\end{figure}
 
@@
 Here we consider how to convert a user's cursor motion into a triangle-mesh representation
-of their stroke.
+of their stroke. This is done in two steps. First, the stroke is rasterized and the
+contour is extracted from the raster. Second, the contour is converted into points and
+boundaries and triangulated.
 
@@
 
-Adam and Steve should we talk about how anti-aliasing + marching squares can achieve sub-pixel
-resolution?
+% Adam and Steve should we talk about how anti-aliasing + marching squares can achieve sub-pixel
+% resolution?
 
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardrendered}
-  \caption{A subfigure}
+  \caption{The rasterized stroke.}
   \label{fig:sub1}
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardpruned}
-  \caption{A subfigure}
+  \caption{The contour of the stroke.}
   \label{fig:sub2}
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardmesh}
-  \caption{A subfigure}
+  \caption{The triangulation of the contour}
   \label{fig:sub2}
 \end{subfigure}
-\caption{A figure with two subfigures}
+\caption{Triangulating a hard stroke.}
 \label{fig:test}
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softrendered}
-  \caption{A subfigure}
+  \caption{The rasterized stroke.}
   \label{fig:sub1}
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softpruned}
-  \caption{A subfigure}
+  \caption{The contour of the stroke.}
   \label{fig:sub2}
@@
   \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softmesh}
-  \caption{A subfigure}
+  \caption{The triangulation of the contour}
   \label{fig:sub2}
 \end{subfigure}
-\caption{A figure with two subfigures}
+\caption{Triangulating a soft stroke.}
 \label{fig:test}
@@
 This representation needs a slightly different procedure. The user's mouse motion
-is capture in two sets of polygons. The first set describes the outer soft stroke and the second set describes
+is captured in two sets of polygons. The first set describes the outer soft stroke and the second set describes
 the inner hard stroke. Next the outer stroke polygons are rendered in 50\% gray and the inner stroke
@@
 stroke where no triangle crosses the boundary between the soft and hard parts
-of the stroke.
+of the stroke. Points on the contour of the inner stroke will have color arcs with the stroke color
+while the points on the contour of the outer stroke will have transparent color arcs. This gives
+all triangles connecting the inner and outer strokes a linear gradient.
 
+\begin{figure}
+    \centering
         \includegraphics[height=1.5in]{images/softstroke}
         \includegraphics[height=1.5in]{images/hardvssoft}
+    \caption{Place Holder.}
+\end{figure}
 
@@
 Once the stroke has been converted to triangles by the above procedure it is necessary to
-composite the stroke onto the canvas. We first consider the canvas model and then how a
-stroke can be composited onto the canvas.
+composite the stroke onto the canvas. This is accomplished in two steps. First, the points
+and boundaries from the triangulation of the stroke are added to the canvas and a new
+triangulation is found. Second, the color arcs for the old and new points are updated.
+We also discuss the challenge intersecting boundaries pose for the algorithm, as well
+as a technique to reduce the number of triangles that are retriangulated on the canvas.
 
@@
 
-Every canvas begins with four points, one at each corner of the screen, and four boundaries, one
-for each edge of the screen. These points are originally part of two big triangles which
+Every canvas begins with a number of points, along the edge of the screen. These
+points have boundaries connecting them. These points are originally part of a number of triangles which
 divide the canvas.
@@
 
+\begin{figure}
+    \centering
         \includegraphics[height=1.5in]{images/canvastriangulation}
+    \caption{Place Holder.}
+\end{figure}
 
@@
 be the first point on each side since this intersection point may be adjacent to other intersection
-points without color data. For both boundaries linearly interpolate the colors in the color arcs. 
+points without color data. For both boundaries the colors in the color arcs are linearly interpolated. 
 This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
@@
 
+\begin{figure}
+    \centering
         \includegraphics[height=1.5in]{images/determiningcolors}
+    \caption{Place Holder.}
+\end{figure}
 
@@
 
+\begin{figure}
+    \centering
         \includegraphics[height=1.5in]{images/findingmodifiedpoints}
+    \caption{Place Holder.}
+\end{figure}
 
 \section{Discussion and Results}
-Our work with this algorithm has lead to observations in three main areas.
+Here we first consider how the complexity of the algorithm grows both in time and geometry.
+The results from this analysis and our experience with TrianglePainter have lead to observations in three main areas.
+
+\subsection{Complexity}
+To find the growth rates of the algorithm we performed two experiments. 
+First, the program drew fifty random strokes without zooming.
+Next, the program drew fifty random strokes with random zooming.
+Geometry grows fastest when many strokes are composited on one another. When not zooming strokes
+must cover the same area of canvas over and over. This
+results in more geometry and therefore takes longer to render. The zooming case explores different
+areas of the canvas and therefore results in less overlapping geometry. This situation should result
+in less complex geometry and therefore take less time.
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.49\textwidth]{graphs/numtriangles}
+    \caption{The blue graph is the non-zooming case. The red graph is the zooming case.}
+    \label{fig:numtriangles}
+\end{figure}
+
+As seen in figure~\ref{fig:numtriangles} the zooming case grows linearly while the non-zooming
+appears to grow quadratically. This is expected since every new stroke has the chance to overlap
+all of the old strokes. This means each new stroke does not just add a constant amount of geometry.
+Due to intersections with previous strokes intersection points are added. This results in a linear
+increase in the amount of geometry. This results
+in a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
+intersection points are rarely added. Instead the geometry increases by a constant which leads to
+a linear growth rate.
+
+\begin{figure}
+    \centering
+        \includegraphics[width=0.49\textwidth]{graphs/cumulativetime}
+    \caption{The blue graph is the non-zooming case. The red graph is the zooming case.}
+    \label{fig:timing}
+\end{figure}
+
+Figure~\ref{fig:timing} gives a similar result for the cumulative time strokes take to render. Both
+the zooming and non-zooming cases grow non-linearly over time. However, the non-zooming case takes
+a lot longer for the reasons described. It's not unexpected that even the zooming case grows non-linearly
+since even though the geometries involved aren't getting very complicated, the geometry of the entire
+canvas is increasing linearly and must be considered even if the entire canvas is not retriangulated.
+
+All this suggests that a simplification is of great importance.
+The case that it is most likely to help is the non-zooming case with complex geometry. When lots of strokes
+overlap there is usually an extra geometry that does not add much detail to the drawing. By removing
+this extra geometry the algorithm should achieve a steady growth rate.
 
@@
 require addressing geometries such as vertices that are too close together. This would
-take more processing time and is something we have yet to explore. However, clearly this
-problem must be solved before the program could be used formally.
+take more processing time and is something we have yet to explore.
 
@@
 a lower level implementation in C++ would be. The drawing of the canvas is quick with
-no lag. However, the actual process to extract a stroke from rasterization to triangle
+no lag. However, the actual process to take a stroke from rasterization to triangle
 representation on the canvas is slow. On a blank canvas strokes take on the order of
@@
 
-\subsection{Scalability}
+\subsection{Simplification}
 No matter how good the implementation, the fact that paintings can get arbitrarily complicated
@@
 without the program slowing to an unusable state.
-
-\subsection{Complexity}
-We now examine the growth of triangles, points, and boundaries on the canvas.
 
@@
 
-The complexity of the triangles increase slightly more than linearly since extra points and 
-boundaries are made when boundaries intersect.
+% The complexity of the triangles increase slightly more than linearly since extra points and 
+% boundaries are made when boundaries intersect.
 
-NOTES:
+% NOTES:
 
--Maybe do the same analysis with soft strokes
+% -Maybe do the same analysis with soft strokes
 
--Look at the time it takes to render these strokes
+% -Look at the time it takes to render these strokes
 
--Then show more robust drawings and the number of triangles, points, and boundaries
+% -Then show more robust drawings and the number of triangles, points, and boundaries
+
+% -Talk about future work in the conclusion
+
+% - Switch 6.3 and 6.4
+\begin{figure*}
+    \centering
+        \includegraphics[width=0.49\textwidth]{images/facezoom}
+        \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
+    \caption{Artist time ~30 minutes. 9030 triangles.}
+\end{figure*}
 
 \section{Conclusion}
+Vector graphics are useful for representing complex images with limited information
+while also enabling easy editing and infinite resolution. Previous work has focused
+on converting raster graphics to vector graphics, unique vector representations, and multi-scale
+raster graphics. However, until now no system
+has enabled artists to create an entire painting in vector graphics using standard brush techniques.
 
-\section*{Acknowledgements}
+TrianglePainter enables a user to paint like they would in a raster graphics program 
+to create vector graphics. The core of TrianglePainter is a triangle mesh representation of the image.
+We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite
+that mesh on the preexisting mesh containing the previous strokes.
+
+The current implementation allows artists to create minimal works. Future work should focus on
+implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
+improvements might focus on creating interesting transforms, or other operations that utilize
+the underlying triangle mesh structure to make the artists' job easier.
+
+\section*{Acknowledgments}
 

### 2014-04-20 template.tex
Add zoom images and macros for TODO notes.
Additional references, small text edits.

@@
 
-\documentclass[conference]{acmsiggraph}
+\documentclass[review]{acmsiggraph}
 \usepackage{subcaption}
+\usepackage{color}
+
 \TOGonlineid{45678}
@@
 \author{Mark D.
-Benjamin\thanks{e-mail:mdbenjam@princeton.edu} \hspace{10 pt} Princeton Unviersity\\ Adam Finkelstein\thanks{e-mail:af@cs.princeton.edu} \hspace{10 pt} Princeton University\\ Stephen Diverdi\thanks{e-mail:diverdi@google.com} \hspace{10 pt} Google}
+Benjamin\thanks{e-mail:mdbenjam@princeton.edu} \hspace{10 pt} Princeton Unviersity\\ Adam Finkelstein\thanks{e-mail:af@cs.princeton.edu} \hspace{10 pt} Princeton University\\ Stephen DiVerdi\thanks{e-mail:diverdi@google.com} \hspace{10 pt} Google}
 \pdfauthor{Mark D. Benjamin, Adam Finkelstein, Stephen Diverdi}
@@
 \keywords{triangle mesh, painting, vector graphics}
+
+\input{macros}
 
@@
 However, traditional computer painting programs take inputs of the same type as the stored data. 
-For example, programs like Photoshop and GIMP
+For example, programs like Adobe Photoshop and GIMP
 allow a user to paint how he would on a physical medium. By letting the mouse act as
@@
 In this paper we introduce TrianglePainter which  
-uses a slightly different representation than traditional raster or vector graphics.
+uses a different representation than traditional raster or vector graphics.
 TrianglePainter uses a triangle mesh to store the data of the painting.
@@
 
+There have been a few projects that enable painting directly with vector data.
+Most similar is the work of DiVerdi et al.~\shortcite{diverdi2012,diverdi2013}, which uses many polygonal paint splats to create a watercolor-like effect.
+These arbitrary polygons are costly to render however, and ``smooth'' effects are only created via many overlapping transparent polygons, which results in a very large amount of geometry.
+\cite{ando2010} also used a dynamic vector representation for their 2D fluid simulation to create marbling patterns stored as B\'{e}zier silhouettes.
+They developed a simulation that adaptively updates the B\'{e}zier to maintain nice contours, but the complexity of the document quickly grows too large to compute interactively.
+
 Another series of works have created entirely new ways of painting. Diffusion curves as specified in \cite{Orzan:2008:DCV:1360612.1360691}
@@
 procedural textures to avoid any fundamental limit on resolution. Although useful,
-generating procedural textures is difficult and does not help most artists easily create.
+generating procedural textures can be difficult for artists.
 
@@
 graphics through standard painting techniques without having to deal with the underlying representation.
-We use the triangle mesh structure described in \cite{10.1109/TVCG.2012.76} to build a painting program
+We start from the triangle mesh structure described in \cite{10.1109/TVCG.2012.76} to build a painting program
 where the triangle mesh is created while the artists lays down strokes.
@@
   \centering
-        \includegraphics[width=0.49\textwidth]{graphs/numtriangles}
-    \caption{The blue graph is the non-zooming case. The red graph is the zooming case.}
+  \begin{subfigure}[b]{0.5\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{graphs/numtriangles}
     \label{fig:numtriangles}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.5\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{graphs/cumulativetime}
+    \label{fig:timing}
+  \end{subfigure}%
+  \vspace{-0.4cm}
+  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting.
+The blue points are the non-zooming case, while the red points are the zooming case.}
 \end{figure}
@@
 a linear growth rate.
-
-\begin{figure}
-    \centering
-        \includegraphics[width=0.49\textwidth]{graphs/cumulativetime}
-    \caption{The blue graph is the non-zooming case. The red graph is the zooming case.}
-    \label{fig:timing}
-\end{figure}
 
@@
         \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
-    \caption{Artist time ~30 minutes. 9030 triangles.}
+    \caption{Artist time ~30 minutes. 9030 triangles. \Steve{Can we add number of strokes?}}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \includegraphics[width=0.195\textwidth]{images/zoom61}
+        \includegraphics[width=0.195\textwidth]{images/zoom62}
+        \includegraphics[width=0.195\textwidth]{images/zoom63}
+        \includegraphics[width=0.195\textwidth]{images/zoom64}
+        \includegraphics[width=0.195\textwidth]{images/zoom65}
+        \includegraphics[width=0.195\textwidth]{images/zoom66}
+        \includegraphics[width=0.195\textwidth]{images/zoom67}
+        \includegraphics[width=0.195\textwidth]{images/zoom68}
+        \includegraphics[width=0.195\textwidth]{images/zoom69}
+        \includegraphics[width=0.195\textwidth]{images/zoom71}
+    \caption{A painting that shows extreme zoom levels.  The final image is at \Steve{zoom level XXX}.}
 \end{figure*}
@@
 
+Omitted for review.
+
 \bibliographystyle{acmsiggraph}

### 2014-04-20 template.tex
More related work, introduction tweaks.

@@
 a brush, a user may color individual pixels on the screen. By contrast, programs like 
-Illustrator and Inkscape let a user paint
+Adobe Illustrator and Inkscape let a user paint
 using geometric primitives such as lines and B\'{e}zier curves. The program then stores
@@
 
-Furthermore, this representation allows painting at any scale. A user can make large strokes
+This representation allows painting at any scale. A user can make large strokes
 while zoomed out and then zoom in to make fine details. All of this can happen without
-loss of quality since the data is stored using triangles, not in pixels.
+loss of quality since the data is stored using triangles instead of pixels, which can represent sharp boundaries at any orientation.
+Furthermore, the use of smooth shading when rendering the triangles allows for vector representation of gradient effects.
+We take advantage of this feature to create soft, airbrush-like strokes, which are usually very difficult to create in vector drawing software.
 
-To transform a user's mouse inputs to a triangle mesh
+To transform a user's mouse inputs to a triangle mesh,
 TrianglePainter uses a combination of rasterization and marching squares to
@@
 
-TrianglePainter reduces the barrier to creating vector graphics. This in turn allows 
+TrianglePainter reduces the barrier to creating vector graphics by enabling a painting-like interface, with support for sharp and smooth edges. This in turn allows 
 artists to focus on the painting and not the way in which they paint, while still 
@@
 They developed a simulation that adaptively updates the B\'{e}zier to maintain nice contours, but the complexity of the document quickly grows too large to compute interactively.
+Finally, \cite{asente2013} implemented a feature in Adobe Illustrator to create contour gradients from shapes, which fill the shapes with small patches of linear gradients to create a bevel-like effect.
+Contour gradients can be used to control opacity to create soft strokes, but each stroke must be stored as a separate (aggregate) object rather than a single flattened canvas representation.
 

### 2014-04-20 template.tex
Edits to section 3.

@@
 \subsection{Triangle Mesh}
-When a user paints he generates points that are stored on the canvas. These points are
+When a user paints, he generates points that are stored on the canvas. These points are
 connected to form a triangle mesh. In every triangle the vertices are colored to either
@@
 multiple triangles can take on different colors in each triangle. Note that although
-the points are persistent, the triangles are not.
+the points are persistent as the canvas is updated, the triangles are not.
+That is, the correct triangulation of the mesh points is not unique, and different equivalent triangulations may be used during the painting process.
 
@@
 boundary edges. A boundary edge connects two points and specifies that no triangle
-can cross it. Like points, boundary edges are persistent. So as new triangles are created they
-may never cross a boundary edges.
+can cross it. Like points, boundary edges are persistent. So as new triangles are created by the triangulation algorithm, they
+may never cross a boundary edges.  This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges in a soft brush stroke.
 
@@
     \centering
-        \includegraphics[width=0.4\textwidth]{images/stroke}
-    \caption{An example triangulation. Blue lines are boundary edges. Black lines form the triangle
+        \includegraphics[width=\columnwidth]{images/stroke}
+    \caption{An example triangulation of a smooth stroke. Blue lines are boundary edges. Black points form the triangle
     mesh with the blue lines.}
+    \label{fig:stroke}
 \end{figure}
@@
 triangles on either side of the boundary. The need for the same point to take on different
-colors in different triangles lead to the development of color arcs.
+colors in different triangles lead to the development of color arcs (see Figure~\ref{fig:arcs}).
 
-A color arc has three components: a start vector, an end vector, and a color. 
+A color arc has three components: a start direction, an end direction, and a color. 
 Determining the color of a vertex in a given triangle requires two steps. First,
@@
 the color arc is found which contains this vector. Containing the vector means one must turn
-clockwise from the starting vector to reach it and counter-clockwise from the ending vector.
+clockwise from the starting direction to reach it and counter-clockwise from the ending direction.
 The point takes on the color of the arc in which the vector from the point to the centroid resides.
 
-The color arcs for every point must satisfy three conditions. First, they must be disjoint. Second,
+A point may have more than one color arc, and they must satisfy three conditions. First, they must be disjoint. Second,
 they must cover all $2\pi$ radians around the point. These first two conditions ensure that for any
-triangle a point will take on exactly one color. Third, the start and end vectors must line up
+triangle a point will take on exactly one color. Third, the start and end vectors must align
 with boundary edges. This condition makes all colors and gradients emanate from boundaries.
@@
     triangles on top and red on bottom with a hard boundary in between.}
+    \label{fig:arcs}
 \end{figure}
@@
 take on a given color based on their color arcs and the centroid of the triangle. If the three
-vertices in the triangle take on the same color value then the triangle will be solid. If instead the 
+vertices in the triangle take on the same color value then the triangle will be flat shaded. If instead the 
 vertices take on different color values then there will be a gradient over the triangle.
 
-Since the conditions on color arcs make colors emanate from boundaries it's easy to create regions
-containing either a solid color or linear gradient. A region will have a solid color if the
+Since the conditions on color arcs make colors emanate from boundaries it is easy to create regions (e.g. a brush stroke boundary)
+containing either a solid color or linear gradient. A region will have a solid color (e.g. a hard brush stroke) if the
 boundaries that surround it all take on the same color. If instead a region is surrounded by
 two boundaries, one which has a given color and the other which has a transparent version of that color
-then there will be a linear gradient in that region between the two boundaries.
+then there will be a linear gradient in that region between the two boundaries (e.g. a soft brush stroke).
 

### 2014-04-20 template.tex
Edits to sections 5, 6, and 7.

@@
 \documentclass[review]{acmsiggraph}
+
 \usepackage{subcaption}
 \usepackage{color}
+
+% Reduce the space between subfigures and subfigure captions.
+\captionsetup[subfigure]{aboveskip=6pt}
+\captionsetup[subfigure]{belowskip=0pt}
 
@@
 \begin{CRcatlist}
-  %\CRcat{I.3.3}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Display Algorithms}
-  %\CRcat{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}{Radiosity};
+  \CRcat{I.3.4}{Computer Graphics}{Graphics Utilities}{Paint systems}
+  \CRcat{I.3.6}{Computer Graphics}{Methodology and Techniques}{Graphics data structures and data types}
 \end{CRcatlist}
@@
 rendered to yield a rasterized version of the stroke.
-Second, to obtain the contours the stroke is rasterized in black and marching squares
-is run to extract iso-contours of 50\% gray. This returns a very dense set of points 
+Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{XXX}
+is used to extract iso-contours of 50\% gray. This returns a very dense set of points 
 which we reduce through a pruning process to
@@
 
-We utilized Jonathan Shewchuk's Triangle library to perform Delaunay triangulations
+We utilized Jonathan Shewchuk's Triangle library~\cite{XXX} to perform constrained Delaunay triangulations
 on the given set of points and boundaries. The library outputs a mesh that
@@
 \subsection{Soft Strokes}
-The above procedure converts a hard brush into a triangle mesh. Yet, a somewhat more interesting
-case is converting a soft brush with linear gradient into a triangle mesh. A soft stroke has two
+The above procedure converts a hard brush stroke into a triangle mesh. A more interesting
+case is converting a soft brush stroke with linear gradient into a triangle mesh. A soft stroke has two
 components, an inner hard stroke and an outer gradient stroke. Since the triangles in the inner and
@@
 Again, using marching squares these contours are extracted and then pruned. The boundaries and points
-derived from the contours are fed into Triangle and a mesh is returned. This mesh represents the soft
-stroke where no triangle crosses the boundary between the soft and hard parts
-of the stroke. Points on the contour of the inner stroke will have color arcs with the stroke color
+derived from the contours are given to Triangle and a mesh is returned. This mesh represents the soft
+stroke where no triangle crosses the boundary between the soft and hard regions
+of the stroke. Points on the contour of the inner stroke will have color arcs with the stroke color,
 while the points on the contour of the outer stroke will have transparent color arcs. This gives
-all triangles connecting the inner and outer strokes a linear gradient.
+all triangles connecting the inner and outer strokes a linear gradient (see Figure~\ref{XXX}).
 
@@
 
-Every canvas begins with a number of points, along the edge of the screen. These
+Every canvas begins with an initial set of points distributed along the edges. These
 points have boundaries connecting them. These points are originally part of a number of triangles which
@@
 
-In the simplest approach the points and boundaries from the new stroke are added
-to the list of points and boundaries already stored. All of this data can be fed into Triangle
+In the simplest approach, the points and boundaries from the new stroke are added
+to the list of points and boundaries already stored on the canvas. All of this data is given to Triangle
 to produce a new mesh. The new mesh contains all of the points of the old and new strokes. Since
-all of the boundaries are stored and fed into Triangle the stroke edges are preserved. This means that
-no triangle can cross the boundary of a stroke. Without this constraint it would be impossible to color the
+all of the boundaries are stored and given to Triangle, the stroke edges are preserved. This means that
+no triangle can cross the boundary of a stroke. Without this constraint, it would be impossible to correctly color the
 triangles since they would straddle color arcs. Once a new mesh is formed the color arcs of the 
@@
 \subsection{Determining Colors}
-The color arcs for new points depend on two values, the color of the stroke and the color
+All color compositing is done with regular RGB alpha blending.  The color arcs for new points depend on two values, the color of the stroke and the color
 of the canvas at that point. The points on a hard stroke will have two color arcs. One
@@
 of the stroke. The color for this arc is just the color of the canvas at that point. The
-points on a soft stroke will have one color arc. The color arc for points on the inner
+points on a soft stroke will each have one color arc. The color arc for points on the inner
 contour is the color of the stroke composited over the color of the canvas at that point.
-The color arc for points on the outer contour is the color of canvas at that point.
+The color arc for points on the outer contour is the color of canvas at that point (because the outer edge of the soft stroke is completely transparent).
 This gives any triangle connecting inside and outside contour points a linear gradient
@@
 
-When Triangle is asked to triangulate a set of points with overlapping boundaries it must
+When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
 insert a point at the intersection, since without it at least one triangle would have to
@@
 To determine the colors of an intersection points requires knowing the colors associated with
-points on the boundaries. For both boundaries travel both directions to the first point with color data. Note this may not
+other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
 be the first point on each side since this intersection point may be adjacent to other intersection
-points without color data. For both boundaries the colors in the color arcs are linearly interpolated. 
+points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
 This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
@@
 \subsection{Finding Modified Points}
-The above procedure has two important drawbacks. First, it processes many points that don't need
+The above procedure has two important drawbacks. First, it processes many points that do not need
 consideration which takes a considerable amount of processing time. Second, since it acts on the whole
@@
 which triangles can remain
-and which triangles need to be included in the re-triangulation. To accomplish this a static grid
+and which triangles need to be included in the re-triangulation (see Figure~\ref{fig:modified}). To accomplish this a static grid
 is generated and all of the cells which contain a triangle from the new stroke are marked. 
@@
     \caption{Place Holder.}
+    \label{fig:modified}
 \end{figure}
@@
 \section{Discussion and Results}
-Here we first consider how the complexity of the algorithm grows both in time and geometry.
-The results from this analysis and our experience with TrianglePainter have lead to observations in three main areas.
+Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face} and \ref{fig:boxes}.  Here we first consider how the complexity of the algorithm grows both in time and geometry.
+The results from this analysis and our experience with TrianglePainter have lead to a number of observations.
 
@@
 First, the program drew fifty random strokes without zooming.
-Next, the program drew fifty random strokes with random zooming.
-Geometry grows fastest when many strokes are composited on one another. When not zooming strokes
+Next, the program drew fifty random strokes with random zoom level for each stroke.
+Geometry grows fastest when many strokes are composited on one another. When not zooming, strokes
 must cover the same area of canvas over and over. This
@@
     \includegraphics[width=\columnwidth]{graphs/numtriangles}
+    \caption{Number of triangles}
     \label{fig:numtriangles}
@@
     \includegraphics[width=\columnwidth]{graphs/cumulativetime}
+    \caption{Rendering time}
     \label{fig:timing}
   \end{subfigure}%
-  \vspace{-0.4cm}
   \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting.
@@
 
-As seen in figure~\ref{fig:numtriangles} the zooming case grows linearly while the non-zooming
+Figure~\ref{fig:numtriangles} shows the zooming case grows linearly while the non-zooming case
 appears to grow quadratically. This is expected since every new stroke has the chance to overlap
@@
 the zooming and non-zooming cases grow non-linearly over time. However, the non-zooming case takes
-a lot longer for the reasons described. It's not unexpected that even the zooming case grows non-linearly
-since even though the geometries involved aren't getting very complicated, the geometry of the entire
-canvas is increasing linearly and must be considered even if the entire canvas is not retriangulated.
+longer for the reasons described. It is not unexpected that even the zooming case grows non-linearly
+since even though the geometries involved are not becoming very complicated, the geometry of the entire
+canvas is increasing linearly and must be considered even if the entire canvas is not re-triangulated.
 
-All this suggests that a simplification is of great importance.
-The case that it is most likely to help is the non-zooming case with complex geometry. When lots of strokes
-overlap there is usually an extra geometry that does not add much detail to the drawing. By removing
+This suggests that canvas simplification is of great importance.
+The case that it is most likely to be helped is the non-zooming case with complex geometry. When lots of strokes
+overlap there is usually extra geometry that does not add much detail to the drawing. By removing
 this extra geometry the algorithm should achieve a steady growth rate.
@@
 \subsection{Stability}
-Something we've worked hard to achieve is program stability. Certain geometries are fatal
-for the Triangle library. The current instability in the program is normally caused by
-providing the library with such difficult geometries. To improve program stability would
-require addressing geometries such as vertices that are too close together. This would
-take more processing time and is something we have yet to explore.
+While the Triangle library is impressively robust, there are still certain pathological geometries that can cause it to fail.
+Specifically, long skinny triangles and vertices that are too close together seem to trigger triangulation failures.
+Detecting these situations and fixing them, or even better modifying the geometry construction to avoid them, would solve this problem.
+For example, identifying if a new vertex is too close to an existing one, and if so, merging them.
+Another example is to subdivide long edges to reduce the incidence of skinny triangles.
+The impact these solutions would have on processing time is something that needs to be explored.
 
@@
 no lag. However, the actual process to take a stroke from rasterization to triangle
-representation on the canvas is slow. On a blank canvas strokes take on the order of
-1 second to process. However, on a canvas with lots of geometry a stroke that covers
-that geometry can take several seconds to render. Some of this time is due to poor
-implementation of the algorithm. However, there are lots of calculations
-to do with the old points affected by the new stroke and the points associated with the
-new stroke itself. Since these geometries can get arbitrarily complicated, these
+representation on the canvas can be slow. On a blank canvas, strokes take less than
+one second to process. However, on a canvas with lots of geometry a stroke that overlaps
+that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
+necessary for the old points affected by the new stroke and the points associated with the
+new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
 calculations can take an arbitrarily long. This brings us to the final point, scalability.
@@
 \subsection{Simplification}
-No matter how good the implementation, the fact that paintings can get arbitrarily complicated
-means there must be some way to simplify the mesh. Although we have yet to implement
-such a simplification algorithm, we have a couple suggestions on how it might be done. In most situations where
-a lot of strokes are composited on top of one another, most of the geometry is redundant. 
-For example, when an opaque stroke is laid down on top of a complex geometry, that complex
+No matter how good the implementation, the fact that paintings can become arbitrarily complicated as more and more strokes are added
+means there must be some way to simplify the mesh. In most situations where
+many strokes overlap one another, most of the geometry is redundant. For example, when an opaque stroke is laid down on top of a complex geometry, that complex
 geometry no longer has any useful information since the outline of the opaque stroke
-completely defines it. Furthermore, even with translucent or feathered strokes, enough
-composited strokes may produce complex geometry that does not add much to the output.
-Some of this geometry can be reduced without noticeable changes in the output. Such a
-simplification algorithm must be implemented to ensure a painter can continue to paint
-without the program slowing to an unusable state.
+completely defines it (all vertices are the same color).  Furthermore, even with translucent or feathered strokes, a large enough number of 
+composited strokes may produce complex geometry that does not add much to the visual appearance of the canvas.
+Some of this geometry can be reduced without noticeable changes.
+To accomplish this, our proposed simplification algorithm would identify a vertex that can be removed via an edge collapse by measuring the change the collapse would cause in the image.
+The change is computed as the magnitude of the color shift and the translation of a boundary.
+If these deltas are below a threshold, then the collapse can proceed.
+This process could be done when a stroke is added to the canvas, or continually in a background thread to avoid increasing apparent latency.
+Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue iterate on a painting without the program becoming too slow.
 
@@
     \caption{Artist time ~30 minutes. 9030 triangles. \Steve{Can we add number of strokes?}}
+    \label{fig:face}
 \end{figure*}
@@
     \caption{A painting that shows extreme zoom levels.  The final image is at \Steve{zoom level XXX}.}
+    \label{fig:boxes}
 \end{figure*}
@@
 \section{Conclusion}
-Vector graphics are useful for representing complex images with limited information
-while also enabling easy editing and infinite resolution. Previous work has focused
-on converting raster graphics to vector graphics, unique vector representations, and multi-scale
-raster graphics. However, until now no system
-has enabled artists to create an entire painting in vector graphics using standard brush techniques.
-
 TrianglePainter enables a user to paint like they would in a raster graphics program 
@@
 
-The current implementation allows artists to create minimal works. Future work should focus on
+The current implementation provides a baseline painting experience. Future work should focus on
 implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
-improvements might focus on creating interesting transforms, or other operations that utilize
+improvements might focus on creating interesting manipulation tools, or other operations that utilize
 the underlying triangle mesh structure to make the artists' job easier.

### 2014-04-21 template.tex
Added new figures to make it look pretty :D

@@
 \begin{abstract}
-Vector graphics have a compact representation, are easy to manipulate, and have infinite resolution.
-Traditionally, vector graphics have required artists to master a new skill in order to effectively utilize
-them. Recent work has converted raster graphics to vector graphics, allowed multi-resolution raster painting,
-and created new ways to represent vector graphics. However, none of these have solved the problem of
-letting an artist use standard and intuitive raster painting techniques to generate vector graphics.
-Our algorithm, TrianglePainter, converts a user's strokes into a triangle mesh vector representation. 
-TrianglePainter lets artists focus on the art, while it handles the vectorization in the background.
+Although vector graphics program have a number of advantages, current vector painting programs
+do not allow an artist to use traditional painting techniques. We propose a new algorithm that translates a
+user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto
+a canvas containing an existing mesh representation of the previous strokes. 
+This representation allows the algorithm to render solid colors and linear
+gradients. It also enables painting at any resolution. This gives artists the opportunity to create complex,
+multi-scale drawings with gradients while avoiding artifacts.
 
@@
 
-This representation allows painting at any scale. A user can make large strokes
+This procedure allows painting at any scale. A user can make large strokes
 while zoomed out and then zoom in to make fine details. All of this can happen without
@@
     \centering
-        \includegraphics[width=0.4\textwidth]{images/colorarcs}
+        \includegraphics[width=0.4\textwidth]{images/colorarcs2}
     \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
@@
 then there will be a linear gradient in that region between the two boundaries (e.g. a soft brush stroke).
+Figure~\ref{fig:softarcs} shows the color arcs on such a soft stroke.
 
@@
     \centering
-        \includegraphics[height=1.5in]{images/softstrokeboundaryedges}
-    \caption{Place Holder.}
+        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
+    \caption{A close look at a gradient stroke. The blue edges are boundaries where
+    the points have gray color arcs as shown. The purple edges are boundaries where
+    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
+    a linear gradient from gray to white.}
+    \label{fig:softarcs}
 \end{figure}
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardrendered}
-  \caption{The rasterized stroke.}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardpruned}
-  \caption{The contour of the stroke.}
-  \label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/hardmesh}
-  \caption{The triangulation of the contour}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{Triangulating a hard stroke.}
-\label{fig:test}
-\end{figure*}
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardrendered}
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softrendered}
 
-\begin{figure*}
-\centering
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softrendered}
-  \caption{The rasterized stroke.}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softpruned}
-  \caption{The contour of the stroke.}
-  \label{fig:sub2}
-\end{subfigure}
-\begin{subfigure}[b]{0.3\textwidth}
-  \centering
-  \includegraphics[width=.9\textwidth]{images/stroke_triangulation/softmesh}
-  \caption{The triangulation of the contour}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{Triangulating a soft stroke.}
-\label{fig:test}
-\end{figure*}
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardpruned}
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softpruned}
+
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardmesh}
+        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softmesh}
+    \caption{Left: A hard stroke is rasterized, the contour is found, and a triangle mesh is generated.
+             Right: A soft stroke is rasterized in two different shades of gray, two contours are found,
+             a triangle mesh is generated with a boundary separating the inner and outer parts of the stroke.}
+    \label{fig:rastertotriangles}
+\end{figure}
 
@@
 while the points on the contour of the outer stroke will have transparent color arcs. This gives
-all triangles connecting the inner and outer strokes a linear gradient (see Figure~\ref{XXX}).
+all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{softarcs}).
+Figure~\ref{fig:rastertotriangles} shows the process of triangulating both hard and soft strokes.
 
-\begin{figure}
-    \centering
-        \includegraphics[height=1.5in]{images/softstroke}
-        \includegraphics[height=1.5in]{images/hardvssoft}
-    \caption{Place Holder.}
-\end{figure}
+% \begin{figure}
+%     \centering
+%         \includegraphics[height=1.5in]{images/softstroke}
+%         \includegraphics[height=1.5in]{images/hardvssoft}
+%     \caption{Place Holder.}
+% \end{figure}
 
@@
 to the list of points and boundaries already stored on the canvas. All of this data is given to Triangle
-to produce a new mesh. The new mesh contains all of the points of the old and new strokes. Since
+to produce a new mesh (see figure~\ref{fig:firststrokes}). The new mesh contains all of the points of the old and new strokes. Since
 all of the boundaries are stored and given to Triangle, the stroke edges are preserved. This means that
@@
     \centering
-        \includegraphics[height=1.5in]{images/canvastriangulation}
-    \caption{Place Holder.}
+        \includegraphics[width=.23\textwidth]{images/emptycanvas}
+        \includegraphics[width=.23\textwidth]{images/canvaswithonestroke}
+    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
+             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
+             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
+    \label{fig:firststrokes}
 \end{figure}
@@
 The colors from the new stroke are composited over the colors from the old stroke to yield four new
-colors, one for each region defined by the boundary intersections.
+colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})
+
+An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.
 
@@
     \centering
-        \includegraphics[height=1.5in]{images/determiningcolors}
-    \caption{Place Holder.}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1new}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2new}
+
+    \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
+             by the the blue stroke and point B is on top of the red stroke. Middle:
+             New color arcs are determined. Both of A's color arcs are composited
+             with the blue stroke which turns red to purple and transparent to blue. Similarly
+             both of B's color arcs are composited on top of the red stroke which turns
+             blue to purple and transparent to red. Two intersection points C and D are
+             added. Their color arcs are determined by examining neighboring points. Right: 
+             A new triangle mesh is generated to preserve boundaries. Note that triangles
+             in the intersection will be purple.}
+    \label{fig:intersectioncolors}
 \end{figure}
+
+\begin{figure*}
+    \centering
+        \includegraphics[width=.32\textwidth]{images/composite1new}
+        \includegraphics[width=.32\textwidth]{images/composite2new}
+        \includegraphics[width=.32\textwidth]{images/composite3new}
+    \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
+             by the the blue stroke and point B is on top of the red stroke. Middle:
+             New color arcs are determined. Both of A's color arcs are composited
+             with the blue stroke which turns red to purple and transparent to blue. Similarly
+             both of B's color arcs are composited on top of the red stroke which turns
+             blue to purple and transparent to red. Two intersection points C and D are
+             added. Their color arcs are determined by examining neighboring points. Right: 
+             A new triangle mesh is generated to preserve boundaries. Note that triangles
+             in the intersection will be purple.}
+    \label{fig:colorarcdeterminer}
+\end{figure*}
 
@@
     \centering
-        \includegraphics[height=1.5in]{images/findingmodifiedpoints}
-    \caption{Place Holder.}
+        \includegraphics[width=.5\textwidth]{images/grid}
+    \caption{An example of how TrianglePainter determines which triangles to save and which to retriangulate. Blue
+    lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
+    the triangles that will be reused, and purple lines define the area that needs to be retriangulated. The purple
+    lines are edges of triangles that intersect the new stroke. Since they interesect the new stroke they are no
+    longer valid and must be retriangulated. All other triangles on the canvas can be reused.}
     \label{fig:modified}
@@
 \section{Discussion and Results}
-Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face} and \ref{fig:boxes}.  Here we first consider how the complexity of the algorithm grows both in time and geometry.
+Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face}, \ref{fig:rainbow}, \ref{fig:stop} and \ref{fig:boxes}.
+Here we first consider how the complexity of the algorithm grows both in time and geometry.
 The results from this analysis and our experience with TrianglePainter have lead to a number of observations.
@@
 appears to grow quadratically. This is expected since every new stroke has the chance to overlap
-all of the old strokes. This means each new stroke does not just add a constant amount of geometry.
-Due to intersections with previous strokes intersection points are added. This results in a linear
-increase in the amount of geometry. This results
-in a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
-intersection points are rarely added. Instead the geometry increases by a constant which leads to
+all of the old strokes. Therefore each new stroke does not just add a constant amount of geometry, but 
+due to intersections with previous strokes extra intersection points are added. This results in a linear
+increase in the amount of geometry per stroke yielding a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
+intersection points are rarely added. In this case the geometry increases by a constant amount per stroke which yields
 a linear growth rate.
@@
 new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
-calculations can take an arbitrarily long. This brings us to the final point, scalability.
+calculations can take an arbitrarily long. This brings us to the final point, simplification.
 
@@
 This process could be done when a stroke is added to the canvas, or continually in a background thread to avoid increasing apparent latency.
-Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue iterate on a painting without the program becoming too slow.
+Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue to iterate on a painting without the program becoming too slow.
 
@@
     \centering
+        \includegraphics[width=0.49\textwidth]{images/rainbow2}
+        % \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
+    \caption{An example of soft strokes}
+    \label{fig:rainbow}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \includegraphics[width=0.3\textwidth]{images/stop1}
+        \includegraphics[width=0.3\textwidth]{images/stop2}
+        \includegraphics[width=0.3\textwidth]{images/stop3}
+    \caption{}
+    \label{fig:stop}
+\end{figure*}
+
+\begin{figure*}
+    \centering
         \includegraphics[width=0.195\textwidth]{images/zoom61}

### 2014-04-21 template.tex
Added the zoom level
Added teaser with rainbow image.

@@
 
-%% \teaser{
-%%   \includegraphics[height=1.5in]{images/sampleteaser}
-%%   \caption{Spring Training 2009, Peoria, AZ.}
-%% }
+\teaser{
+  \setlength{\w}{0.33\textwidth}
+  \centering
+  \includegraphics[width=\w]{images/rainbow3}
+  \includegraphics[width=\w]{images/rainbow3_tri}
+  \includegraphics[width=\w]{images/rainbow3_zoom3}
+  \caption{A vector painting made in TrianglePaint that exhibits smooth gradients painted by soft brushes, as well as infinitely sharp boundaries painted by hard brushes, together in a single flat representation.
+The triangle mesh is shown in the middle, and the right image is zoomed in on the bottom of the letter B, to show the sharp boundaries and smooth gradients in detail.}
+  \label{fig:teaser}
+}
 
@@
         \includegraphics[width=0.195\textwidth]{images/zoom71}
-    \caption{A painting that shows extreme zoom levels.  The final image is at \Steve{zoom level XXX}.}
+    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 93,000.}
     \label{fig:boxes}

### 2014-04-21 template.tex
Moving figures, added missing references.

@@
 
+\begin{figure}
+    \centering
+        \includegraphics[width=\columnwidth]{images/stroke}
+    \caption{An example triangulation of a smooth stroke. Blue lines are boundary edges. Black points form the triangle
+    mesh with the blue lines.}
+    \label{fig:stroke}
+\end{figure}
+
 There have been a few projects that enable painting directly with vector data.
@@
 
+\begin{figure}
+    \centering
+        \includegraphics[width=0.4\textwidth]{images/colorarcs2}
+    \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
+    mesh with the boundary edges. Each point is given a color arc which defines its color in that range.
+    Note that the points in the middle row have different colors on top and bottom. This gives blue
+    triangles on top and red on bottom with a hard boundary in between.}
+    \label{fig:arcs}
+\end{figure}
+
 \section{The Canvas Model}
@@
 
-\begin{figure}
-    \centering
-        \includegraphics[width=\columnwidth]{images/stroke}
-    \caption{An example triangulation of a smooth stroke. Blue lines are boundary edges. Black points form the triangle
-    mesh with the blue lines.}
-    \label{fig:stroke}
-\end{figure}
-
 \subsection{Color Arcs}
@@
 with boundary edges. This condition makes all colors and gradients emanate from boundaries.
-
-\begin{figure}
-    \centering
-        \includegraphics[width=0.4\textwidth]{images/colorarcs2}
-    \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
-    mesh with the boundary edges. Each point is given a color arc which defines its color in that range.
-    Note that the points in the middle row have different colors on top and bottom. This gives blue
-    triangles on top and red on bottom with a hard boundary in between.}
-    \label{fig:arcs}
-\end{figure}
 
@@
 rendered to yield a rasterized version of the stroke.
-Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{XXX}
+Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{lorensen1987}
 is used to extract iso-contours of 50\% gray. This returns a very dense set of points 
@@
 
-We utilized Jonathan Shewchuk's Triangle library~\cite{XXX} to perform constrained Delaunay triangulations
+We utilized Jonathan Shewchuk's Triangle library~\cite{shewchuk2002} to perform constrained Delaunay triangulations
 on the given set of points and boundaries. The library outputs a mesh that
@@
 while the points on the contour of the outer stroke will have transparent color arcs. This gives
-all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{softarcs}).
+all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{fig:softarcs}).
 Figure~\ref{fig:rastertotriangles} shows the process of triangulating both hard and soft strokes.
@@
 
-\subsection{Intersection Points}
-
-When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
-insert a point at the intersection, since without it at least one triangle would have to
-cross a boundary. These points are introduced with no color information, but the final colors for 
-all of the other points in the mesh have already been determined.
-
-An intersection point lies on two boundaries and therefore has four boundary edges emanating from it. 
-To determine the colors of an intersection points requires knowing the colors associated with
-other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
-be the first point on each side since this intersection point may be adjacent to other intersection
-points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
-This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
-The colors from the new stroke are composited over the colors from the old stroke to yield four new
-colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})
-
-An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.
-
 \begin{figure}
@@
 
+\subsection{Intersection Points}
+
+When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
+insert a point at the intersection, since without it at least one triangle would have to
+cross a boundary. These points are introduced with no color information, but the final colors for 
+all of the other points in the mesh have already been determined.
+
+An intersection point lies on two boundaries and therefore has four boundary edges emanating from it. 
+To determine the colors of an intersection points requires knowing the colors associated with
+other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
+be the first point on each side since this intersection point may be adjacent to other intersection
+points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
+This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
+The colors from the new stroke are composited over the colors from the old stroke to yield four new
+colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})
+
+An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.
+
 \subsection{Finding Modified Points}
@@
 
-\begin{figure}
-    \centering
-        \includegraphics[width=.5\textwidth]{images/grid}
-    \caption{An example of how TrianglePainter determines which triangles to save and which to retriangulate. Blue
-    lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
-    the triangles that will be reused, and purple lines define the area that needs to be retriangulated. The purple
-    lines are edges of triangles that intersect the new stroke. Since they interesect the new stroke they are no
-    longer valid and must be retriangulated. All other triangles on the canvas can be reused.}
-    \label{fig:modified}
-\end{figure}
-
 \section{Discussion and Results}
-Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face}, \ref{fig:rainbow}, \ref{fig:stop} and \ref{fig:boxes}.
+Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.
 Here we first consider how the complexity of the algorithm grows both in time and geometry.
@@
 in less complex geometry and therefore take less time.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[b]{0.5\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{graphs/numtriangles}
-    \caption{Number of triangles}
-    \label{fig:numtriangles}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{graphs/cumulativetime}
-    \caption{Rendering time}
-    \label{fig:timing}
-  \end{subfigure}%
-  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting.
-The blue points are the non-zooming case, while the red points are the zooming case.}
-\end{figure}
 
@@
 
+\begin{figure}
+    \centering
+        \includegraphics[width=\columnwidth]{images/grid}
+    \caption{An example of how TrianglePainter determines which triangles to save and which to retriangulate. Blue
+    lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
+    the triangles that will be reused, and purple lines define the area that needs to be retriangulated. The purple
+    lines are edges of triangles that intersect the new stroke. Since they interesect the new stroke they are no
+    longer valid and must be retriangulated. All other triangles on the canvas can be reused.}
+    \label{fig:modified}
+\end{figure}
+
 \subsection{Performance}
@@
 Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue to iterate on a painting without the program becoming too slow.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{0.5\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{graphs/numtriangles}
+    \caption{Number of triangles}
+    \label{fig:numtriangles}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.5\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{graphs/cumulativetime}
+    \caption{Rendering time}
+    \label{fig:timing}
+  \end{subfigure}%
+  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting.
+The blue points are the non-zooming case, while the red points are the zooming case.}
+\end{figure}
 
@@
     \label{fig:face}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-        \includegraphics[width=0.49\textwidth]{images/rainbow2}
-        % \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
-    \caption{An example of soft strokes}
-    \label{fig:rainbow}
 \end{figure*}

### 2014-04-21 paper.tex
cleanup file names
foo

--- 2014-04-21 template.tex
@@
 
-\title{A Triangle Mesh Approach to Painting}
+\title{Painting with Triangles}
 
@@
   \includegraphics[width=\w]{images/rainbow3_zoom3}
-  \caption{A vector painting made in TrianglePaint that exhibits smooth gradients painted by soft brushes, as well as infinitely sharp boundaries painted by hard brushes, together in a single flat representation.
-The triangle mesh is shown in the middle, and the right image is zoomed in on the bottom of the letter B, to show the sharp boundaries and smooth gradients in detail.}
+  \caption{\emph{Left:} A vector painting made in TrianglePaint exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
+\emph{Middle:} The painting is represented by a planar triangle mesh.
+\emph{Right:} Closeup on the bottom of the letter B highlights both the sharp boundaries on the white region and smooth gradient from blue to green in the background.}
   \label{fig:teaser}
@@
 \begin{abstract}
-Although vector graphics program have a number of advantages, current vector painting programs
-do not allow an artist to use traditional painting techniques. We propose a new algorithm that translates a
+Although vector graphics offer a number of benefits, conventional vector painting programs
+do not allow an artist to use traditional (digital) painting techniques. 
+% \note{really?}
+We propose a new algorithm that translates a
 user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto
@@
 Furthermore, the use of smooth shading when rendering the triangles allows for vector representation of gradient effects.
-We take advantage of this feature to create soft, airbrush-like strokes, which are usually very difficult to create in vector drawing software.
+We take advantage of this feature to create soft, airbrush-like strokes, which are difficult to create in vector drawing software.
 
@@
     \centering
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardrendered}
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softrendered}
+    \setlength{\w}{0.45\columnwidth}
 
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardpruned}
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softpruned}
+        \includegraphics[width=\w]{images/stroke_triangulation/hardrendered}
+        \includegraphics[width=\w]{images/stroke_triangulation/softrendered}
 
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/hardmesh}
-        \includegraphics[width=.2\textwidth]{images/stroke_triangulation/softmesh}
+        \includegraphics[width=\w]{images/stroke_triangulation/hardpruned}
+        \includegraphics[width=\w]{images/stroke_triangulation/softpruned}
+
+        \includegraphics[width=\w]{images/stroke_triangulation/hardmesh}
+        \includegraphics[width=\w]{images/stroke_triangulation/softmesh}
     \caption{Left: A hard stroke is rasterized, the contour is found, and a triangle mesh is generated.
@@
         \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
-    \caption{Artist time ~30 minutes. 9030 triangles. \Steve{Can we add number of strokes?}}
+    \caption{A drawing by an artist that took ~30 minutes. 9030 triangles.} % \Steve{Can we add number of strokes?}}
     \label{fig:face}
@@
         \includegraphics[width=0.3\textwidth]{images/stop3}
-    \caption{}
+    \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
     \label{fig:stop}
@@
 \bibliographystyle{acmsiggraph}
-\bibliography{template}
+\bibliography{paper}
 \end{document}

### 2014-04-21 paper.tex
Updated figures and captions
fixing citations

@@
 Previous work has focused on converting images into vector graphics to take advantage of the
-efficient storage, easy editing, and infinite resolution it provides. For example \cite{lecot:ARD:2006}
+efficient storage, easy editing, and infinite resolution it provides. For example the method of
+Lecot and L\'{e}vy~\shortcite{lecot:ARD:2006}
 minimizes an energy function to segment an image into a number of regions that are bounded
@@
 
-Other work has proposed novel vector graphic data structures. For example \cite{Frisken:2000:ASD:344779.344899}
-defines Adaptively Sampled Distance Fields (ADF). ADFs specify a signed distance function to a surface.
-Rendering the shape requires sampling the function to determine whether or not a pixel is on the surface.
+Other work has proposed novel vector graphic data structures. For example 
+Frisken~\etal~\shortcite{Frisken:2000:ASD:344779.344899}
+introduced Adaptively Sampled Distance Fields (ADFs), which specify a signed distance function to a surface in 3D
+(or a curve in the plane).
+%
+Rendering the shape requires sampling the function to determine whether or not a pixel is on the shape.
 A quadtree accompanies the distance function to specify where higher sampling rates are necessary. \cite{Bremer:2001:VCM}
@@
     \centering
-        \includegraphics[width=0.4\textwidth]{images/colorarcs2}
+        \includegraphics[width=0.4\textwidth]{images/colorarcsfinal}
     \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
@@
     \centering
-        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1new}
-        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2new}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}
 
-    \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
-             by the the blue stroke and point B is on top of the red stroke. Middle:
-             New color arcs are determined. Both of A's color arcs are composited
-             with the blue stroke which turns red to purple and transparent to blue. Similarly
-             both of B's color arcs are composited on top of the red stroke which turns
-             blue to purple and transparent to red. Two intersection points C and D are
-             added. Their color arcs are determined by examining neighboring points. Right: 
-             A new triangle mesh is generated to preserve boundaries. Note that triangles
-             in the intersection will be purple.}
+    \caption{Left: A red stroke is composited over a blue stroke. An intersection
+            point is inserted in the center and four color arcs must be determined. Right:
+            The four colors are resolved as the four combinations of the red stroke's color
+            arcs composited over the blue stroke's color arcs.}
     \label{fig:intersectioncolors}
@@
     \centering
-        \includegraphics[width=.32\textwidth]{images/composite1new}
-        \includegraphics[width=.32\textwidth]{images/composite2new}
-        \includegraphics[width=.32\textwidth]{images/composite3new}
+        \includegraphics[width=.32\textwidth]{images/composite21}
+        \includegraphics[width=.32\textwidth]{images/composite22}
+        \includegraphics[width=.32\textwidth]{images/composite23}
     \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
@@
     \centering
-        \includegraphics[width=0.3\textwidth]{images/stop1}
-        \includegraphics[width=0.3\textwidth]{images/stop2}
-        \includegraphics[width=0.3\textwidth]{images/stop3}
+        \includegraphics[width=0.3\textwidth]{images/stop12}
+        \includegraphics[width=0.3\textwidth]{images/stop22}
+        \includegraphics[width=0.3\textwidth]{images/stop32}
     \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}

### 2014-04-21 paper.tex
Changed citations to shortcite

@@
 minimizes an energy function to segment an image into a number of regions that are bounded
-by cubic splines and filled with solid colors or gradients. A different approach by \cite{Lai:2009:ATG:1531326.1531391} 
-automatically creates gradient meshes with support for holes. Yet another example is \cite{10.1109/TVCG.2012.76} 
+by cubic splines and filled with solid colors or gradients. A different approach by Lai~\etal~\shortcite{Lai:2009:ATG:1531326.1531391} 
+automatically creates gradient meshes with support for holes. Yet another example is the work of Liao~\etal~\shortcite{10.1109/TVCG.2012.76} 
 which converts an image into a triangle mesh. All three of these can convert images into
@@
 Rendering the shape requires sampling the function to determine whether or not a pixel is on the shape.
-A quadtree accompanies the distance function to specify where higher sampling rates are necessary. \cite{Bremer:2001:VCM}
+A quadtree accompanies the distance function to specify where higher sampling rates are necessary. The method of Bremer~\etal~\shortcite{Bremer:2001:VCM}
 further describes how ADFs can be created and used. However, it is unclear how ADFs might be used in a general
@@
 These arbitrary polygons are costly to render however, and ``smooth'' effects are only created via many overlapping transparent polygons, which results in a very large amount of geometry.
-\cite{ando2010} also used a dynamic vector representation for their 2D fluid simulation to create marbling patterns stored as B\'{e}zier silhouettes.
+The method of Ando and Tsuruno~\shortcite{ando2010} also uses a dynamic vector representation for their 2D fluid simulation to create marbling patterns stored as B\'{e}zier silhouettes.
 They developed a simulation that adaptively updates the B\'{e}zier to maintain nice contours, but the complexity of the document quickly grows too large to compute interactively.
-Finally, \cite{asente2013} implemented a feature in Adobe Illustrator to create contour gradients from shapes, which fill the shapes with small patches of linear gradients to create a bevel-like effect.
+Finally, Asente and Carr~\shortcite{asente2013} implemented a feature in Adobe Illustrator to create contour gradients from shapes, which fill the shapes with small patches of linear gradients to create a bevel-like effect.
 Contour gradients can be used to control opacity to create soft strokes, but each stroke must be stored as a separate (aggregate) object rather than a single flattened canvas representation.
 
-Another series of works have created entirely new ways of painting. Diffusion curves as specified in \cite{Orzan:2008:DCV:1360612.1360691}
-describe a new painting technique where the artists specify edges and colors for those edges. Their system
+Another series of works have created entirely new ways of painting. Orzan~\etal~\shortcite{Orzan:2008:DCV:1360612.1360691}
+specified Diffusion curves which
+describe a new painting technique where artists specify edges and colors for those edges. Their system
 then solves a Poisson equation to diffuse the colors between boundary conditions specified at the edges.
-Similarly \cite{McCann:2008:RGP:1360612.1360692}
+Similarly the method of McCann and Pollard~\shortcite{McCann:2008:RGP:1360612.1360692}
 lets artists paint in the gradient domain. Both of these approaches yield fantastic works of art, yet
@@
 
-Work has also been done in multi-resolution painting using raster graphics. \cite{Berman:1994:MPC:192161.192181}
+Work has also been done in multi-resolution painting using raster graphics. The method of Berman~\etal~\shortcite{Berman:1994:MPC:192161.192181}
 stores image information in a quadtree so that different parts of the image can have different levels of detail.
@@
 there are still limitations. For example a stroke painted at coarse detail will still look blurry when
-zoomed in. The strokes are limited to the resolution they are painted at. \cite{Carr:2004:PD:1186562.1015809}
+zoomed in. The strokes are limited to the resolution they are painted at. Similarly the method of Carr and Hart~\shortcite{Carr:2004:PD:1186562.1015809}
 also allows the development of multi-resolution images using rasters, and as such suffers from the
-same limitations. Another multi-resolution approach by \cite{Perlin:1995:LPP:218380.218437} uses
+same limitations. Another multi-resolution approach by Perlin and Velho~\shortcite{Perlin:1995:LPP:218380.218437} uses
 procedural textures to avoid any fundamental limit on resolution. Although useful,
@@
 graphics through standard painting techniques without having to deal with the underlying representation.
-We start from the triangle mesh structure described in \cite{10.1109/TVCG.2012.76} to build a painting program
+We start from the triangle mesh structure described by Liao~\etal~\shortcite{10.1109/TVCG.2012.76} to build a painting program
 where the triangle mesh is created while the artists lays down strokes.

### 2014-04-21 paper.tex
foo

--- 2014-04-21 paper.tex
@@
         \includegraphics[width=\columnwidth]{images/stroke}
-    \caption{An example triangulation of a smooth stroke. Blue lines are boundary edges. Black points form the triangle
-    mesh with the blue lines.}
+    \caption{Example smooth stroke triangulation. Blue segments are boundary edges, and small blue squares denote
+    mesh vertices.}
     \label{fig:stroke}
@@
 the points are persistent as the canvas is updated, the triangles are not.  
-That is, the correct triangulation of the mesh points is not unique, and different equivalent triangulations may be used during the painting process.
+That is, triangulation of the mesh points is not unique, and different, equivalent triangulations may 
+be used throughout the painting process.
+Thus, we refer to them as ``mesh points'' rather than ``vertices'' (which implies a particular graph
+structure).
 
 \subsection{Boundary Edges}
-Although the triangles may change over time, they must abide by constraints known as
-boundary edges. A boundary edge connects two points and specifies that no triangle
-can cross it. Like points, boundary edges are persistent. So as new triangles are created by the triangulation algorithm, they
-may never cross a boundary edges.  This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges in a soft brush stroke.
+Although the triangles may change over time, they must abide by constraints we call
+\emph{boundary edges}, each of which connects a pair of points and specifies that no triangle
+can cross it. Like mesh points, boundary edges are persistent. 
+%
+So as new triangles are created by the triangulation algorithm, they
+may never cross a boundary edge.  
+This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft brush stroke.
 

### 2014-04-21 paper.tex
foo

@@
 artists to focus on the painting and not the way in which they paint, while still 
-getting the benefits of vector graphics.
+getting the benefits of vector graphics. We demonstrate our method by creating digital paintings with a 
+range of effects and resolutions, including zoom factors of up to 500,000:1 with an effective resolution
+of an image with more than $10^{17}$ pixels.
 
-\section{Related Works}
+\section{Related Work}
 
@@
 by cubic splines and filled with solid colors or gradients. A different approach by Lai~\etal~\shortcite{Lai:2009:ATG:1531326.1531391} 
-automatically creates gradient meshes with support for holes. Yet another example is the work of Liao~\etal~\shortcite{10.1109/TVCG.2012.76} 
-which converts an image into a triangle mesh. All three of these can convert images into
-a vector representation, yet none of them allow a user to create a painting in the vector medium.
+automatically creates gradient meshes with support for holes. 
+Our work builds on the representation introduced by
+Liao~\etal~\shortcite{10.1109/TVCG.2012.76}, who showed that
+a photo can be converted into a triangle mesh.
+They demonstrated that natural images can be concisely represented
+by triangles, using mesh simplification, and that
+various image editing operations can be applied in this representation,
+including an abstraction technique similar to that of 
+DeCarlo and Santella~\shortcite{DeCarlo:2002:SAP:566654.566650}.
+In principle, one could create a digital painting in a conventional
+program and then use one of these approaches to convert to a vector
+representation. In contrast our paper shows how to paint directly
+in this representation, offering several advantages including
+efficiency and the ability to handle imagery with enormously
+higher effective resolution.
+% All three of these can convert images into
+% a vector representation, yet none of them allow a user to create a painting in the vector medium.
 
@@
 Our work builds on these findings. However, our goal is to enable artists to create vector
-graphics through standard painting techniques without having to deal with the underlying representation.
+graphics through standard painting techniques without having to consider the underlying representation.
+\ignorethis{
 We start from the triangle mesh structure described by Liao~\etal~\shortcite{10.1109/TVCG.2012.76} to build a painting program
 where the triangle mesh is created while the artists lays down strokes. 
+}
 

### 2014-04-21 paper.tex
Updated a figure

--- 2014-04-21 paper.tex
@@
         \includegraphics[width=\columnwidth]{images/stroke}
-    \caption{Example smooth stroke triangulation. Blue segments are boundary edges, and small blue squares denote
-    mesh vertices.}
+    \caption{An example triangulation of a smooth stroke. Blue lines are boundary edges. Black points form the triangle
+    mesh with the blue lines.}
     \label{fig:stroke}
@@
 the points are persistent as the canvas is updated, the triangles are not.
-That is, triangulation of the mesh points is not unique, and different, equivalent triangulations may 
-be used throughout the painting process.
-Thus, we refer to them as ``mesh points'' rather than ``vertices'' (which implies a particular graph
-structure).
+That is, the correct triangulation of the mesh points is not unique, and different equivalent triangulations may be used during the painting process.
 
 \subsection{Boundary Edges}
-Although the triangles may change over time, they must abide by constraints we call
-\emph{boundary edges}, each of which connects a pair of points and specifies that no triangle
-can cross it. Like mesh points, boundary edges are persistent. 
-%
-So as new triangles are created by the triangulation algorithm, they
-may never cross a boundary edge.  
-This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft brush stroke.
+Although the triangles may change over time, they must abide by constraints known as
+boundary edges. A boundary edge connects two points and specifies that no triangle
+can cross it. Like points, boundary edges are persistent. So as new triangles are created by the triangulation algorithm, they
+may never cross a boundary edges.  This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges in a soft brush stroke.
 
@@
         \includegraphics[width=0.195\textwidth]{images/zoom69}
-        \includegraphics[width=0.195\textwidth]{images/zoom71}
-    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 93,000.}
+        \includegraphics[width=0.195\textwidth]{images/zoom72}
+    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000.}
     \label{fig:boxes}

### 2014-04-21 paper.tex
Minor tweak to remove dangling line.
foo

--- 2014-04-21 paper.tex
@@
 
+\ignorethis{
 \begin{CRcatlist}
@@
 \end{CRcatlist}
+}
 
@@
 getting the benefits of vector graphics. We demonstrate our method by creating digital paintings with a 
-range of effects and resolutions, including zoom factors of up to 500,000:1 with an effective resolution
-of an image with more than $10^{17}$ pixels.
+range of effects and resolutions, including zoom factors of up to 500,000:1 which implies an effective 
+resolution equivalent to an image with more than $10^{17}$ pixels.
 

### 2014-04-21 paper.tex
Added paper submission number

@@
 
-\TOGonlineid{45678}
+\TOGonlineid{33}
 \TOGvolume{0}
@@
   \includegraphics[width=\w]{images/rainbow3_zoom3}
-  \caption{\emph{Left:} A vector painting made in TrianglePaint exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
+  \caption{\emph{Left:} A vector painting made in the system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
 \emph{Middle:} The painting is represented by a planar triangle mesh.
@@
 
-In this paper we introduce TrianglePainter which  
+In this paper we introduce a system which  
 uses a different representation than traditional raster or vector graphics.
-TrianglePainter uses a triangle mesh to store the data of the painting.
+The system uses a triangle mesh to store the data of the painting.
 Since triangles are a geometric primitive used in vector graphics this approach achieves the
@@
 To transform a user's mouse inputs to a triangle mesh,
-TrianglePainter uses a combination of rasterization and marching squares to
+the system uses a combination of rasterization and marching squares to
 find stroke contours. From the contour it performs a Delaunay triangulation
@@
 
-TrianglePainter reduces the barrier to creating vector graphics by enabling a painting-like interface, with support for sharp and smooth edges. This in turn allows 
+The system reduces the barrier to creating vector graphics by enabling a painting-like interface, with support for sharp and smooth edges. This in turn allows 
 artists to focus on the painting and not the way in which they paint, while still 
@@
 may never cross a boundary edge.  
-This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft brush stroke.
+This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft  stroke.
 
@@
 \section{Discussion and Results}
-Paintings made with TrianglePainter can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.
+Paintings made with the system can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.
 Here we first consider how the complexity of the algorithm grows both in time and geometry.
-The results from this analysis and our experience with TrianglePainter have lead to a number of observations.
+The results from this analysis and our experience with the system have lead to a number of observations.
 
@@
         \includegraphics[width=\columnwidth]{images/grid}
-    \caption{An example of how TrianglePainter determines which triangles to save and which to retriangulate. Blue
+    \caption{An example of how the system determines which triangles to save and which to retriangulate. Blue
     lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
@@
 \section{Conclusion}
-TrianglePainter enables a user to paint like they would in a raster graphics program 
-to create vector graphics. The core of TrianglePainter is a triangle mesh representation of the image.
+The system enables a user to paint like they would in a raster graphics program 
+to create vector graphics. The core of the system is a triangle mesh representation of the image.
 We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite

### 2014-04-21 paper.tex
foo

--- 2014-04-21 paper.tex
@@
 A quadtree accompanies the distance function to specify where higher sampling rates are necessary. The method of Bremer~\etal~\shortcite{Bremer:2001:VCM}
-further describes how ADFs can be created and used. However, it is unclear how ADFs might be used in a general
+further describes how ADFs can be created and used. However, it remains unclear how ADFs might be used in a general
 painting program where strokes with solid colors and gradients are composited on top of one another.
@@
 
-Another series of works have created entirely new ways of painting. Orzan~\etal~\shortcite{Orzan:2008:DCV:1360612.1360691}
+Another line of research has explored entirely new ways of painting.
+Orzan~\etal~\shortcite{Orzan:2008:DCV:1360612.1360691}
 specified Diffusion curves which
@@
 Similarly the method of McCann and Pollard~\shortcite{McCann:2008:RGP:1360612.1360692}
-lets artists paint in the gradient domain. Both of these approaches yield fantastic works of art, yet
-they are really new mediums for artists to explore rather than a vectorized version of conventional
-painting techniques.
+lets artists paint in the gradient domain. Both of these approaches yield creative new aesthetic ranges 
+beyond what is straightforward in the traditional pixel-based painting model. Nevertheless,
+they represent a new painting metaphor for the artist, rather than a vectorized version of the conventional
+digital painting approach familiar to artists.
 
@@
 may never cross a boundary edge.  
-This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft brush stroke.
+This partially restricts the set of valid triangulations of the points.  Figure~\ref{fig:stroke} shows boundary edges for a soft  stroke.
+
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
+    \caption{A close look at a gradient stroke. The blue edges are boundaries where
+    the points have gray color arcs as shown. The purple edges are boundaries where
+    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
+    a linear gradient from gray to white.}
+    \label{fig:softarcs}
+\end{figure}
 
@@
 Figure~\ref{fig:softarcs} shows the color arcs on such a soft stroke.
-
-\begin{figure}
-    \centering
-        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
-    \caption{A close look at a gradient stroke. The blue edges are boundaries where
-    the points have gray color arcs as shown. The purple edges are boundaries where
-    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
-    a linear gradient from gray to white.}
-    \label{fig:softarcs}
-\end{figure}
 

### 2014-04-21 paper.tex

--- 2014-04-21 paper.tex
@@
 
+\begin{figure}[t]
+    \centering
+        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
+    \caption{A close look at a gradient stroke. The blue edges are boundaries where
+    the points have gray color arcs as shown. The purple edges are boundaries where
+    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
+    a linear gradient from gray to white.}
+    \label{fig:softarcs}
+\end{figure}
+
 \subsection{Color Arcs}

### 2014-04-21 paper.tex
Changed figures
foo

@@
 Orzan~\etal~\shortcite{Orzan:2008:DCV:1360612.1360691}
-specified Diffusion curves which
-describe a new painting technique where artists specify edges and colors for those edges. Their system
+introduced \emph{diffusion curves} --
+a new painting technique where artists specify edges and colors for those edges. Their system
 then solves a Poisson equation to diffuse the colors between boundary conditions specified at the edges.
@@
         \includegraphics[width=0.4\textwidth]{images/colorarcsfinal}
-    \caption{An example of color arcs. The orange lines are boundary edges. Black lines form the triangle
-    mesh with the boundary edges. Each point is given a color arc which defines its color in that range.
-    Note that the points in the middle row have different colors on top and bottom. This gives blue
-    triangles on top and red on bottom with a hard boundary in between.}
+    \caption{Color arcs. Orange lines are boundary edges, and together with black lines form the triangle
+    mesh. Each point is given more or more color arcs, 
+    which stores the color of triangle corners in that angular range.
+    Points in the middle row have red arcs on top and blue below,
+    yielding triangles with those colors, separated by a hard boundary.}
     \label{fig:arcs}
@@
 A triangle mesh must contain all the points and stay
-within the boundaries. For example, a concave contour should not contain triangles inside the concavity since such triangles
-would be outside of the boundary.
-Furthermore, these boundaries will later be crucial when compositing a stroke onto a canvas.
-
-We utilized Jonathan Shewchuk's Triangle library~\cite{shewchuk2002} to perform constrained Delaunay triangulations
-on the given set of points and boundaries. The library outputs a mesh that
+within the boundaries. 
+For example, a concave contour like the outline of the letter C should not contain 
+triangles inside the concavity, since such triangles would be outside of the boundary.
+These boundaries will prove to be crucial when compositing a stroke onto the canvas.
+%
+We perform constrained Delaunay triangulation on the given set of points and boundaries using
+Shewchuk's implementation~\shortcite{shewchuk2002}, and the output is a mesh that
 approximately represents the drawn stroke.
@@
 \subsection{Soft Strokes}
-The above procedure converts a hard brush stroke into a triangle mesh. A more interesting
+The above procedure converts a hard brush stroke into a triangle mesh. A more challenging
 case is converting a soft brush stroke with linear gradient into a triangle mesh. A soft stroke has two

### 2014-04-21 paper.tex
foo

@@
 Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{lorensen1987}
-is used to extract iso-contours of 50\% gray. This returns a very dense set of points 
+is used to extract iso-contours of 50\% gray. This returns a very dense set of points,
 which we reduce through a pruning process to
 get a sparse but accurate representation of the contour.
+The pruning procedure walks around the contour to form an approximating polygon $P$,
+greedily eliding vertices whenever they lie within a small threshold distance of $P$.
 

### 2014-04-22 paper.tex
Updated boxes
Fix tilde.
move figure.
Increase figure 12 size.
foo

@@
 
+% uncomment one the following two lines, first one showing pictures
 \documentclass[review]{acmsiggraph}
+%\documentclass[review,draft]{acmsiggraph}
 
@@
 \usepackage{color}
+
+% comment out this line to get figures; uncomment for fast latex
+%\usepackage[draft]{graphicx}
+\usepackage{graphicx}
 
@@
 \TOGcodeURL{}
+
+\setlength{\fboxsep}{0pt}
+\setlength{\fboxrule}{1pt}
 
@@
   \includegraphics[width=\w]{images/rainbow3_zoom3}
-  \caption{\emph{Left:} A vector painting made in the system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
+  \caption{\emph{Left:} A vector painting made in our system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
 \emph{Middle:} The painting is represented by a planar triangle mesh.
-\emph{Right:} Closeup on the bottom of the letter B highlights both the sharp boundaries on the white region and smooth gradient from blue to green in the background.}
+\emph{Right:} Closeup showing the bottom of the letter B reveals both the sharp boundaries on the white region and a smooth gradient from blue to green in the background.}
   \label{fig:teaser}
@@
 Although vector graphics offer a number of benefits, conventional vector painting programs
-do not allow an artist to use traditional (digital) painting techniques. 
-% \note{really?}
+offer only limited support for the traditional painting metaphor.
 We propose a new algorithm that translates a
-user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto
-a canvas containing an existing mesh representation of the previous strokes. 
+user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. 
 This representation allows the algorithm to render solid colors and linear
-gradients. It also enables painting at any resolution. This gives artists the opportunity to create complex,
-multi-scale drawings with gradients while avoiding artifacts.
+gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.
 
@@
 To transform a user's mouse inputs to a triangle mesh,
-the system uses a combination of rasterization and marching squares to
+our system uses a combination of rasterization and marching squares to
 find stroke contours. From the contour it performs a Delaunay triangulation
@@
     \centering
-        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
+        \fbox{\includegraphics[width=0.45\textwidth]{images/gradientarcs}}
     \caption{A close look at a gradient stroke. The blue edges are boundaries where
@@
 \subsection{Converting Strokes to Contours}
+
 A set of contours define the outline of a stroke. For simply connected strokes, those without holes, 
@@
 \subsection{Converting Contours to Triangles}
-The contour contains two important pieces of information. First, it describes all of the points of the stroke.
-Second, segments connecting adjacent points describe the boundaries of the stroke. 
+\label{sec:contours}
+
+Now that we have the contour we need to create a triangulation for the stroke. 
+First we consider the case of a hard brush, which contains a single contour (\fig{fig:rastertotriangles}-left).
+Points on the contour describe the shape of the stroke, while the adjoining segments
+describe its boundary.
 A triangle mesh must contain all the points and stay
@@
 Again, using marching squares these contours are extracted and then pruned. The boundaries and points
-derived from the contours are given to Triangle and a mesh is returned. This mesh represents the soft
+derived from the contours are triangulated as in \sect{sec:contours}.
+The resulting mesh represents the soft
 stroke where no triangle crosses the boundary between the soft and hard regions
@@
 while the points on the contour of the outer stroke will have transparent color arcs. This gives
-all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{fig:softarcs}).
-Figure~\ref{fig:rastertotriangles} shows the process of triangulating both hard and soft strokes.
+all triangles connecting the inner and outer strokes a linear gradient as seen in Figure~\ref{fig:softarcs}.
+Figure~\ref{fig:rastertotriangles}-right shows the process for triangulating soft strokes.
 
@@
 
+\begin{figure}
+    \centering
+        \fbox{\includegraphics[width=.22\textwidth]{images/emptycanvas}}
+        \fbox{\includegraphics[width=.22\textwidth]{images/canvaswithonestroke}}
+    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
+             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
+             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
+    \label{fig:firststrokes}
+\end{figure}
+
 \subsection{Adding Points to the Canvas}
@@
 In the simplest approach, the points and boundaries from the new stroke are added
-to the list of points and boundaries already stored on the canvas. All of this data is given to Triangle
-to produce a new mesh (see figure~\ref{fig:firststrokes}). The new mesh contains all of the points of the old and new strokes. Since
-all of the boundaries are stored and given to Triangle, the stroke edges are preserved. This means that
-no triangle can cross the boundary of a stroke. Without this constraint, it would be impossible to correctly color the
-triangles since they would straddle color arcs. Once a new mesh is formed the color arcs of the 
-new and old points must be computed.
-
-\begin{figure}
-    \centering
-        \includegraphics[width=.23\textwidth]{images/emptycanvas}
-        \includegraphics[width=.23\textwidth]{images/canvaswithonestroke}
-    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
-             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
-             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
-    \label{fig:firststrokes}
-\end{figure}
+to the list of points and boundaries already stored on the canvas. 
+%
+These points and boundaries are all triangulated using constrained Delaunay triangulation
+as in \sect{sec:contours}, 
+to produce a new mesh (Figure~\ref{fig:firststrokes}).
+%
+The new mesh contains all of the points of the old and new strokes. 
+Since boundary edges form constraints in the triangulation, 
+stroke edges are preserved. This means that
+no triangle will cross the boundary of a stroke. 
+Without this constraint, it would be impossible to correctly color the
+triangles since they would straddle color arcs.
 
 \subsection{Determining Colors}
+
+Once a new mesh is formed the color arcs of the 
+new and old points must be computed.
+%
 All color compositing is done with regular RGB alpha blending.  The color arcs for new points depend on two values, the color of the stroke and the color
 of the canvas at that point. The points on a hard stroke will have two color arcs. One
-arc describes the inside of the stroke. The color for this arc is the color of the stroke
+arc describes the inside of the stroke, which takes the color of the stroke
 composited over the color of the canvas at that point. The other arc describes the outside
-of the stroke. The color for this arc is just the color of the canvas at that point. The
+of the stroke, which is just the color of the canvas at that point. The
 points on a soft stroke will each have one color arc. The color arc for points on the inner
@@
 The color arc for points on the outer contour is the color of canvas at that point (because the outer edge of the soft stroke is completely transparent).
-This gives any triangle connecting inside and outside contour points a linear gradient
-from the stroke color composited over the canvas color to the canvas color.
+This arrangement produces a linear gradient across the triangles separating the inner
+and outer contours.
 
-The color arcs for old points remain the same unless part of the new stroke covers it.
+The color arcs for old points remain unchanged, unless part of the new stroke covers it.
 In this case the colors in each arc must be replaced with the stroke color composited over
@@
         \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
+        \hspace{.2in}
         \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}
@@
 
-When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
-insert a point at the intersection, since without it at least one triangle would have to
-cross a boundary. These points are introduced with no color information, but the final colors for 
+When the constrained Delaunay triangulation is computed
+for a set of points with intersecting boundary edges, it must
+insert a new point at the intersection of the edges
+(otherwise at least one triangle would have to
+cross a boundary). These points are introduced with no color information, 
+whereas the final colors for 
 all of the other points in the mesh have already been determined.
@@
 To determine the colors of an intersection points requires knowing the colors associated with
-other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
-be the first point on each side since this intersection point may be adjacent to other intersection
+other points on the boundaries. For both boundaries, 
+we search both directions to the first point with color data. Note this may not
+be the first point on each side, since this intersection point may be adjacent to 
+\emph{other} intersection
 points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
@@
 The colors from the new stroke are composited over the colors from the old stroke to yield four new
-colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})
-
-An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.
+colors, one for each region defined by the boundary intersections (see Figure~\ref{fig:intersectioncolors}).
+% XXX
+For an example of how color arcs are determined for both intersection and regular points see Figure~\ref{fig:colorarcdeterminer}.
 
 \subsection{Finding Modified Points}
-The above procedure has two important drawbacks. First, it processes many points that do not need
-consideration which takes a considerable amount of processing time. Second, since it acts on the whole
-canvas, local changes can have global effects. To avoid this problem the algorithm determines
-which triangles can remain
+
+As described so far, our retriangulation procedure has two substantial drawbacks. 
+%
+First, it processes many points that do not need consideration which is inefficient. 
+Second, since it acts on the whole
+canvas, local modifications may have global effects. 
+%
+Instead our approach is to determine
+which triangles can remain unchanged
 and which triangles need to be included in the re-triangulation (see Figure~\ref{fig:modified}). To accomplish this a static grid
-is generated and all of the cells which contain a triangle from the new stroke are marked. 
+is generated and all of the cells that contain a triangle from the new stroke are marked. 
 The set of triangles from the old triangulation that intersect these grid squares are found.
@@
 these points and edges gives the new geometry for the modified area and has no effect on any
-other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused. 
-The new triangulation combined with the old triangles give a new triangle mesh representation of the entire canvas.
+other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused,
+and their union with the new triangle mesh represents the new canvas.
 
 \section{Discussion and Results}
-Paintings made with the system can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.
+Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.
 Here we first consider how the complexity of the algorithm grows both in time and geometry.
@@
 due to intersections with previous strokes extra intersection points are added. This results in a linear
-increase in the amount of geometry per stroke yielding a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
-intersection points are rarely added. In this case the geometry increases by a constant amount per stroke which yields
-a linear growth rate.
+increase in the amount of geometry per new stroke, yielding a quadratic growth rate overall. 
+However, in the zooming case each stroke rarely overlaps previous strokes, and therefore
+intersection points are rarely added. In this case the geometry increases by a constant amount per stroke, which yields a linear growth rate.
 
@@
 \subsection{Stability}
-While the Triangle library is impressively robust, there are still certain pathological geometries that can cause it to fail.
+
+While Shewchuk's implementation~\shortcite{shewchuk2002} of Delaunay triangulation is impressively robust, there are still certain pathological geometries that can cause it to fail.
 Specifically, long skinny triangles and vertices that are too close together seem to trigger triangulation failures.
@@
     \centering
-        \includegraphics[width=\columnwidth]{images/grid}
-    \caption{An example of how the system determines which triangles to save and which to retriangulate. Blue
+        \fbox{\includegraphics[width=.45\textwidth]{images/grid}}
+    \caption{An example of how our system determines which triangles to save and which to retriangulate. Blue
     lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
@@
     \centering
-    \includegraphics[width=\columnwidth]{graphs/numtriangles}
+    % \includegraphics[width=\columnwidth]{graphs/numtriangles}
+    \includegraphics[height=1.2in]{graphs/numtriangles}
     \caption{Number of triangles}
@@
     \centering
-    \includegraphics[width=\columnwidth]{graphs/cumulativetime}
+    % \includegraphics[width=\columnwidth]{graphs/cumulativetime}
+    \includegraphics[height=1.2in]{graphs/cumulativetime}
     \caption{Rendering time}
@@
     \centering
-        \includegraphics[width=0.49\textwidth]{images/facezoom}
-        \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
-    \caption{A drawing by an artist that took ~30 minutes. 9030 triangles.} % \Steve{Can we add number of strokes?}}
+        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoom}}
+        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
+    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
     \label{fig:face}
@@
     \centering
-        \includegraphics[width=0.3\textwidth]{images/stop12}
-        \includegraphics[width=0.3\textwidth]{images/stop22}
-        \includegraphics[width=0.3\textwidth]{images/stop32}
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
     \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
@@
         \includegraphics[width=0.195\textwidth]{images/zoom72}
-    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000.}
+    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1, for an effective resolution equivalent to an image with more than $10^{17}$ pixels.}
     \label{fig:boxes}
@@
 \section{Conclusion}
-The system enables a user to paint like they would in a raster graphics program 
-to create vector graphics. The core of the system is a triangle mesh representation of the image.
-We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite
-that mesh on the preexisting mesh containing the previous strokes.
+Our system enables a user to paint like they would in a raster graphics program 
+to create vector graphics. The core element is a triangle mesh representation of the painting.
+We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.
 
-The current implementation provides a baseline painting experience. Future work should focus on
+There are a number of intersting areas for future work, including:
+
+%\noindent
+%{\bf Simplification.}
+%Our prototype implementation provides a baseline painting experience, in
+
+Future work should focus on
 implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
@@
 
+\ignorethis{
 \section*{Acknowledgments}
-
 Omitted for review.
+}
 

### 2014-04-22 paper.tex
foo

@@
 % uncomment one the following two lines, first one showing pictures
-\documentclass[review]{acmsiggraph}
-%\documentclass[review,draft]{acmsiggraph}
+%\documentclass[review]{acmsiggraph}
+\documentclass[review,draft]{acmsiggraph}
 
@@
 % comment out this line to get figures; uncomment for fast latex
-%\usepackage[draft]{graphicx}
-\usepackage{graphicx}
+\usepackage[draft]{graphicx}
+%\usepackage{graphicx}
 
@@
 new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
-calculations can take an arbitrarily long. This brings us to the final point, simplification.
+calculations can take an arbitrarily long.
+% This brings us to the final point, simplification.
 
 \subsection{Simplification}
+\label{sec:simplify}
 No matter how good the implementation, the fact that paintings can become arbitrarily complicated as more and more strokes are added
@@
 Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue to iterate on a painting without the program becoming too slow.
+%
+Finally we note that Liao~\etal~\shortcite{10.1109/TVCG.2012.76} demonstrated that this kind of simplification can be effective at reducing mesh complexity, at least for natural imagery.
+%
+In their case they were working on an offline process for the entire image, whereas our
+problem needs to be interactive on mouse-up (or as a background thread) but fortunately
+only needs to operate on regions of the mesh that have changed during recent painting.
 
@@
 
-There are a number of intersting areas for future work, including:
+Aside from the simplification challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:
 
-%\noindent
-%{\bf Simplification.}
-%Our prototype implementation provides a baseline painting experience, in
+\noindent
+{\bf Curved boundaries.}
+In our implementation, curve boundaries are represented as polylines. We believe it would be relatively straightforward to modify these boundaries to be represented by higher-order curves, 
+for example a Catmull-Rom spline passing through the vertices of the mesh. This would offer
+higher-quality paths for the same sampling rate, but would require subdividing the path at
+rendering time based on the image resolution. Today such operations can typically be 
+performed on graphics hardware with little impact on runtime performance.
 
-Future work should focus on
-implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
-improvements might focus on creating interesting manipulation tools, or other operations that utilize
-the underlying triangle mesh structure to make the artists' job easier.
+\noindent
+{\bf Geometric operations.}
+Because the underlying representation is vector in nature, it should robustly support warping and
+other geometric operations. It should also be easy to provide other kinds of useful tools to the 
+artist like a ``smoothing brush'' which could locally adjust the geometry and/or the colors to soften a region.
+Finally, it would be interesting to offer this kind of painting interface on non-planar canvasses, for example allowing an artist to paint directly onto a manifold representing a shape in 3D.
 

### 2014-04-22 paper.tex
foo

@@
 % uncomment one the following two lines, first one showing pictures
-%\documentclass[review]{acmsiggraph}
-\documentclass[review,draft]{acmsiggraph}
+\documentclass[review]{acmsiggraph}
+%\documentclass[review,draft]{acmsiggraph}
 
@@
 % comment out this line to get figures; uncomment for fast latex
-\usepackage[draft]{graphicx}
-%\usepackage{graphicx}
+\usepackage{graphicx}
 
@@
 
-\section{Discussion and Results}
+\section{Results and Discussion}
 Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.  
-Here we first consider how the complexity of the algorithm grows both in time and geometry.
+%
+These figures contain drawings in a range of styles. The images shown in \fig{fig:boxes}
+reveal a drawing with an extreme zoom range up to 524,000:1, 
+for an effective resolution equivalent to an image with more than $10^{17}$ pixels.
+At this zoom level our render begins to reach the limits of floating point precision
+for single floats; this could be ameliorated by switching to double-precision.
+
+Next we  consider how the complexity of the algorithm grows both in time and geometry.
 The results from this analysis and our experience with the system have lead to a number of observations.
@@
         \includegraphics[width=0.195\textwidth]{images/zoom72}
-    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1, for an effective resolution equivalent to an image with more than $10^{17}$ pixels.}
+    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1.}
     \label{fig:boxes}

### 2014-04-22 paper.tex
Added the eye and apple images

--- 2014-04-22 paper.tex
@@
 
-% uncomment one the following two lines, first one showing pictures
 \documentclass[review]{acmsiggraph}
-%\documentclass[review,draft]{acmsiggraph}
 
@@
 \usepackage{color}
-
-% comment out this line to get figures; uncomment for fast latex
-%\usepackage[draft]{graphicx}
-\usepackage{graphicx}
 
@@
   \includegraphics[width=\w]{images/rainbow3_zoom3}
-  \caption{\emph{Left:} A vector painting made in our system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
+  \caption{\emph{Left:} A vector painting made in the system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes.
 \emph{Middle:} The painting is represented by a planar triangle mesh.
-\emph{Right:} Closeup showing the bottom of the letter B reveals both the sharp boundaries on the white region and a smooth gradient from blue to green in the background.}
+\emph{Right:} Closeup on the bottom of the letter B highlights both the sharp boundaries on the white region and smooth gradient from blue to green in the background.}
   \label{fig:teaser}
@@
 Although vector graphics offer a number of benefits, conventional vector painting programs
-offer only limited support for the traditional painting metaphor.
+do not allow an artist to use traditional (digital) painting techniques. 
+% \note{really?}
 We propose a new algorithm that translates a
-user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. 
+user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto
+a canvas containing an existing mesh representation of the previous strokes. 
 This representation allows the algorithm to render solid colors and linear
-gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.
+gradients. It also enables painting at any resolution. This gives artists the opportunity to create complex,
+multi-scale drawings with gradients while avoiding artifacts.
 
@@
 To transform a user's mouse inputs to a triangle mesh,
-our system uses a combination of rasterization and marching squares to
+the system uses a combination of rasterization and marching squares to
 find stroke contours. From the contour it performs a Delaunay triangulation
@@
 \subsection{Converting Strokes to Contours}
-
 A set of contours define the outline of a stroke. For simply connected strokes, those without holes, 
@@
 \subsection{Converting Contours to Triangles}
-\label{sec:contours}
-
-Now that we have the contour we need to create a triangulation for the stroke. 
-First we consider the case of a hard brush, which contains a single contour (\fig{fig:rastertotriangles}-left).
-Points on the contour describe the shape of the stroke, while the adjoining segments
-describe its boundary.
+The contour contains two important pieces of information. First, it describes all of the points of the stroke.
+Second, segments connecting adjacent points describe the boundaries of the stroke. 
 A triangle mesh must contain all the points and stay
@@
 Again, using marching squares these contours are extracted and then pruned. The boundaries and points
-derived from the contours are triangulated as in \sect{sec:contours}.
-The resulting mesh represents the soft
+derived from the contours are given to Triangle and a mesh is returned. This mesh represents the soft
 stroke where no triangle crosses the boundary between the soft and hard regions
@@
 while the points on the contour of the outer stroke will have transparent color arcs. This gives
-all triangles connecting the inner and outer strokes a linear gradient as seen in Figure~\ref{fig:softarcs}.
-Figure~\ref{fig:rastertotriangles}-right shows the process for triangulating soft strokes.
+all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{fig:softarcs}).
+Figure~\ref{fig:rastertotriangles} shows the process of triangulating both hard and soft strokes.
 
@@
 
+\subsection{Adding Points to the Canvas}
+
+Every canvas begins with an initial set of points distributed along the edges. These
+points have boundaries connecting them. These points are originally part of a number of triangles which
+divide the canvas.
+
+In the simplest approach, the points and boundaries from the new stroke are added
+to the list of points and boundaries already stored on the canvas. All of this data is given to Triangle
+to produce a new mesh (see figure~\ref{fig:firststrokes}). The new mesh contains all of the points of the old and new strokes. Since
+all of the boundaries are stored and given to Triangle, the stroke edges are preserved. This means that
+no triangle can cross the boundary of a stroke. Without this constraint, it would be impossible to correctly color the
+triangles since they would straddle color arcs. Once a new mesh is formed the color arcs of the 
+new and old points must be computed.
+
 \begin{figure}
@@
 
-\subsection{Adding Points to the Canvas}
-
-Every canvas begins with an initial set of points distributed along the edges. These
-points have boundaries connecting them. These points are originally part of a number of triangles which
-divide the canvas.
-
-In the simplest approach, the points and boundaries from the new stroke are added
-to the list of points and boundaries already stored on the canvas. 
-%
-These points and boundaries are all triangulated using constrained Delaunay triangulation
-as in \sect{sec:contours}, 
-to produce a new mesh (Figure~\ref{fig:firststrokes}).
-%
-The new mesh contains all of the points of the old and new strokes. 
-Since boundary edges form constraints in the triangulation, 
-stroke edges are preserved. This means that
-no triangle will cross the boundary of a stroke. 
-Without this constraint, it would be impossible to correctly color the
-triangles since they would straddle color arcs.
-
 \subsection{Determining Colors}
-
-Once a new mesh is formed the color arcs of the 
-new and old points must be computed.
-%
 All color compositing is done with regular RGB alpha blending.  The color arcs for new points depend on two values, the color of the stroke and the color
 of the canvas at that point. The points on a hard stroke will have two color arcs. One
-arc describes the inside of the stroke, which takes the color of the stroke
+arc describes the inside of the stroke. The color for this arc is the color of the stroke
 composited over the color of the canvas at that point. The other arc describes the outside
-of the stroke, which is just the color of the canvas at that point. The
+of the stroke. The color for this arc is just the color of the canvas at that point. The
 points on a soft stroke will each have one color arc. The color arc for points on the inner
@@
 The color arc for points on the outer contour is the color of canvas at that point (because the outer edge of the soft stroke is completely transparent).
-This arrangement produces a linear gradient across the triangles separating the inner
-and outer contours.
+This gives any triangle connecting inside and outside contour points a linear gradient
+from the stroke color composited over the canvas color to the canvas color.
 
-The color arcs for old points remain unchanged, unless part of the new stroke covers it.
+The color arcs for old points remain the same unless part of the new stroke covers it.
 In this case the colors in each arc must be replaced with the stroke color composited over
@@
 
-When the constrained Delaunay triangulation is computed
-for a set of points with intersecting boundary edges, it must
-insert a new point at the intersection of the edges
-(otherwise at least one triangle would have to
-cross a boundary). These points are introduced with no color information, 
-whereas the final colors for 
+When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
+insert a point at the intersection, since without it at least one triangle would have to
+cross a boundary. These points are introduced with no color information, but the final colors for 
 all of the other points in the mesh have already been determined.
@@
 To determine the colors of an intersection points requires knowing the colors associated with
-other points on the boundaries. For both boundaries, 
-we search both directions to the first point with color data. Note this may not
-be the first point on each side, since this intersection point may be adjacent to 
-\emph{other} intersection
+other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
+be the first point on each side since this intersection point may be adjacent to other intersection
 points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
@@
 The colors from the new stroke are composited over the colors from the old stroke to yield four new
-colors, one for each region defined by the boundary intersections (see Figure~\ref{fig:intersectioncolors}).
-% XXX
-For an example of how color arcs are determined for both intersection and regular points see Figure~\ref{fig:colorarcdeterminer}.
+colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})
+
+An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.
 
 \subsection{Finding Modified Points}
-
-As described so far, our retriangulation procedure has two substantial drawbacks. 
-%
-First, it processes many points that do not need consideration which is inefficient. 
-Second, since it acts on the whole
-canvas, local modifications may have global effects. 
-%
-Instead our approach is to determine
-which triangles can remain unchanged
+The above procedure has two important drawbacks. First, it processes many points that do not need
+consideration which takes a considerable amount of processing time. Second, since it acts on the whole
+canvas, local changes can have global effects. To avoid this problem the algorithm determines
+which triangles can remain
 and which triangles need to be included in the re-triangulation (see Figure~\ref{fig:modified}). To accomplish this a static grid
-is generated and all of the cells that contain a triangle from the new stroke are marked. 
+is generated and all of the cells which contain a triangle from the new stroke are marked. 
 The set of triangles from the old triangulation that intersect these grid squares are found.
@@
 these points and edges gives the new geometry for the modified area and has no effect on any
-other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused,
-and their union with the new triangle mesh represents the new canvas.
+other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused. 
+The new triangulation combined with the old triangles give a new triangle mesh representation of the entire canvas.
 
 \section{Discussion and Results}
-Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.
+Paintings made with the system can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.
 Here we first consider how the complexity of the algorithm grows both in time and geometry.
@@
 due to intersections with previous strokes extra intersection points are added. This results in a linear
-increase in the amount of geometry per new stroke, yielding a quadratic growth rate overall. 
-However, in the zooming case each stroke rarely overlaps previous strokes, and therefore
-intersection points are rarely added. In this case the geometry increases by a constant amount per stroke, which yields a linear growth rate.
+increase in the amount of geometry per stroke yielding a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
+intersection points are rarely added. In this case the geometry increases by a constant amount per stroke which yields
+a linear growth rate.
 
@@
 \subsection{Stability}
-
-While Shewchuk's implementation~\shortcite{shewchuk2002} of Delaunay triangulation is impressively robust, there are still certain pathological geometries that can cause it to fail.
+While the Triangle library is impressively robust, there are still certain pathological geometries that can cause it to fail.
 Specifically, long skinny triangles and vertices that are too close together seem to trigger triangulation failures.
@@
         \fbox{\includegraphics[width=.45\textwidth]{images/grid}}
-    \caption{An example of how our system determines which triangles to save and which to retriangulate. Blue
+    \caption{An example of how the system determines which triangles to save and which to retriangulate. Blue
     lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
@@
         \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
-    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
+    \caption{A drawing by an artist that took 30 minutes. 9030 triangles. Approximately 70 strokes.}
     \label{fig:face}
@@
     \centering
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
+        \fbox{\includegraphics[width=0.3\textwidth]{images/stop12}}
+        \fbox{\includegraphics[width=0.3\textwidth]{images/stop22}}
+        \fbox{\includegraphics[width=0.3\textwidth]{images/stop32}}
     \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
+    \label{fig:stop}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \fbox{\includegraphics[width=0.45\textwidth]{images/apple}}
+        \fbox{\includegraphics[width=0.45\textwidth]{images/eye}}
+    \caption{Two more example paintings exploring different styles.}
     \label{fig:stop}
@@
         \includegraphics[width=0.195\textwidth]{images/zoom72}
-    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1, for an effective resolution equivalent to an image with more than $10^{17}$ pixels.}
+    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000.}
     \label{fig:boxes}
@@
 \section{Conclusion}
-Our system enables a user to paint like they would in a raster graphics program 
-to create vector graphics. The core element is a triangle mesh representation of the painting.
-We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.
+The system enables a user to paint like they would in a raster graphics program 
+to create vector graphics. The core of the system is a triangle mesh representation of the image.
+We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite
+that mesh on the preexisting mesh containing the previous strokes.
 
-There are a number of intersting areas for future work, including:
-
-%\noindent
-%{\bf Simplification.}
-%Our prototype implementation provides a baseline painting experience, in
-
-Future work should focus on
+The current implementation provides a baseline painting experience. Future work should focus on
 implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
@@
 
-\ignorethis{
 \section*{Acknowledgments}
+
 Omitted for review.
-}
 

### 2014-04-22 paper.tex

--- 2014-04-22 paper.tex
@@
         \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
-    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
+    \caption{A drawing by an artist that took about 30 minutes, using about 70 strokes, resulting in 9030 triangles.}
     \label{fig:face}
@@
     \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
+    \label{fig:stop}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \fbox{\includegraphics[width=0.45\textwidth]{images/apple}}
+        \fbox{\includegraphics[width=0.45\textwidth]{images/eye}}
+    \caption{Two more example paintings exploring different styles.}
     \label{fig:stop}

### 2014-04-22 paper.tex
figure rearrangement.

--- 2014-04-22 paper.tex
@@
 \begin{abstract}
-Although vector graphics offer a number of benefits, conventional vector painting programs
-offer only limited support for the traditional painting metaphor.
-We propose a new algorithm that translates a
-user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. 
-This representation allows the algorithm to render solid colors and linear
-gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.
-
+Although vector graphics offer a number of benefits, conventional vector painting programs offer only limited support for the traditional painting metaphor.
+We propose a new algorithm that translates a user's mouse motion into a triangle mesh representation.
+This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes.
+This representation allows the algorithm to render solid colors and linear gradients.
+It also enables painting at any resolution.
+This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.
 \end{abstract}
@@
 \section{Results and Discussion}
-Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.  
+Paintings made with our system in a range of visual styles are shown in Figures~\ref{fig:teaser} and \ref{fig:face}--\ref{fig:boxes}.  
 %
-These figures contain drawings in a range of styles. The images shown in \fig{fig:boxes}
+The images shown in \fig{fig:boxes}
 reveal a drawing with an extreme zoom range up to 524,000:1, 
 for an effective resolution equivalent to an image with more than $10^{17}$ pixels.
-At this zoom level our render begins to reach the limits of floating point precision
-for single floats; this could be ameliorated by switching to double-precision.
+At this zoom level our renderer begins to suffer from artifacts at 
+the limits of single floating point precision; 
+this could be ameliorated by switching to double-precision.
 
@@
 The current implementation uses Python and PyOpenGL. Python's performance is slower than
-a lower level implementation in C++ would be. The drawing of the canvas is quick with
+a lower level implementation in C++ would be. The preview rasterization of a new stroke on 
+the canvas is interactive with
 no lag. However, the actual process to take a stroke from rasterization to triangle
@@
         \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
-    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
+    \caption{This drawing uses about 70 strokes, resulting in 9030 triangles.}
     \label{fig:face}
@@
         \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
-    \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
+    \caption{This painting shows soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
     \label{fig:stop}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \fbox{\includegraphics[height=0.333\textwidth]{images/apple}}
+		\hspace*{0.5in}
+        \fbox{\includegraphics[height=0.333\textwidth]{images/eye}}
+    \caption{Two more example paintings exploring different styles.}
+    \label{fig:apple-eye}
 \end{figure*}
@@
         \includegraphics[width=0.195\textwidth]{images/zoom72}
-    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1.}
+    \caption{This single painting shows an extreme zoom; the final closeup image is at zoom level 524,000:1.}
     \label{fig:boxes}
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 Our system enables a user to paint like they would in a raster graphics program 
@@
 
-Aside from the simplification challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:
+Aside from the {\bf simplification} challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:
 

### 2014-04-22 paper.tex
figure rearrange
Made a few minor changes to sentences that had extra words etc.

@@
     \caption{Color arcs. Orange lines are boundary edges, and together with black lines form the triangle
-    mesh. Each point is given more or more color arcs, 
-    which stores the color of triangle corners in that angular range.
+    mesh. Each point is given one or more color arcs 
+    which store the color of triangle corners in that angular range.
     Points in the middle row have red arcs on top and blue below,
@@
     the points have gray color arcs as shown. The purple edges are boundaries where
-    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
+    the points have white color arcs. This gives triangles connecting the two boundaries
     a linear gradient from gray to white.}
@@
 \section{Results and Discussion}
-Paintings made with our system in a range of visual styles are shown in Figures~\ref{fig:teaser} and \ref{fig:face}--\ref{fig:boxes}.  
+Paintings made with our system in a range of visual styles are shown in Figures~\ref{fig:teaser} and \ref{fig:face}--\ref{fig:stop}.  
 %
@@
 Figure~\ref{fig:numtriangles} shows the zooming case grows linearly while the non-zooming case
-appears to grow quadratically. This is expected since every new stroke has the chance to overlap
+appears to grow quadratically. This is expected since in the non-zooming case every new stroke has the chance to overlap
 all of the old strokes. Therefore each new stroke does not just add a constant amount of geometry, but 
@@
 new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
-calculations can take an arbitrarily long.
+calculations can take arbitrarily long.
 % This brings us to the final point, simplification.
@@
     \centering
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
-        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
-    \caption{This painting shows soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
-    \label{fig:stop}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-        \fbox{\includegraphics[height=0.333\textwidth]{images/apple}}
-		\hspace*{0.5in}
-        \fbox{\includegraphics[height=0.333\textwidth]{images/eye}}
-    \caption{Two more example paintings exploring different styles.}
-    \label{fig:apple-eye}
-\end{figure*}
-
-\begin{figure*}
-    \centering
         \includegraphics[width=0.195\textwidth]{images/zoom61}
@@
 
+\begin{figure*}
+    \centering
+        \fbox{\includegraphics[height=0.333\textwidth]{images/apple}}
+		\hspace*{0.5in}
+        \fbox{\includegraphics[height=0.333\textwidth]{images/eye}}
+    \caption{Two more example paintings exploring different styles.}
+    \label{fig:apple-eye}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
+        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
+    \caption{This painting shows soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
+    \label{fig:stop}
+\end{figure*}
+
 \section{Conclusion and Future Work}
@@
 to create vector graphics. The core element is a triangle mesh representation of the painting.
-We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.
+We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite it onto the preexisting mesh containing the previous strokes.
 

### 2014-06-05 paper.tex
foo
Changed graph labels and made changes based on reviewers suggestions.

@@
 
-Work has also been done in multi-resolution painting using raster graphics. The method of Berman~\etal~\shortcite{Berman:1994:MPC:192161.192181}
+Work has also been done in multi-resolution painting using raster graphics. \modified{Early work with Pad++ by Bederson and Hollan~\shortcite{Bederson:1994:PZG:192426.192435} 
+explored the possibility of having an infinite resolution canvas. However, rather than a general painting program 
+this work is focused on creating an interface builder for developers.}
+The method of Berman~\etal~\shortcite{Berman:1994:MPC:192161.192181}
 stores image information in a quadtree so that different parts of the image can have different levels of detail.
@@
 also allows the development of multi-resolution images using rasters, and as such suffers from the
-same limitations. Another multi-resolution approach by Perlin and Velho~\shortcite{Perlin:1995:LPP:218380.218437} uses
+same limitations. \modified{One way to reduce blurriness in raster images of sharp features is proposed by Ramanarayanan~\etal~\shortcite{ramanarayanan2004feature}. 
+By seprately storing where important boundaries are, their system can avoid excessive interpolation. However, such a method requires
+creating such a data structure with those stored boundaries.} Another multi-resolution approach by Perlin and Velho~\shortcite{Perlin:1995:LPP:218380.218437} uses
 procedural textures to avoid any fundamental limit on resolution. Although useful,
@@
 First, the user's mouse movements are captured by a set of polygons which can be
-rendered to yield a rasterized version of the stroke.
+rendered to yield a rasterized version of the stroke. \modified{This rasterization happens at the resolution of the current window. Therefore
+whatever the artist sees on the screen is the basis for the vectorization.}
 Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{lorensen1987}
@@
 The pruning procedure walks around the contour to form an approximating polygon $P$,
-greedily eliding vertices whenever they lie within a small threshold distance of $P$.
+greedily eliding vertices whenever they lie within a small threshold distance of $P$. \modified{We use
+a threshold relative to zoom of 0.5 which was determined to work well empirically.}
 
@@
 the limits of single floating point precision; 
-this could be ameliorated by switching to double-precision.
+this could be ameliorated by switching to double-precision. \modified{Traditional vector
+graphics software such as Adobe Illustrator caps zooming at a ratio of 64:1. Although this
+is not an inherent limitation, there is currently not a way to use such vector graphics programs
+to draw using a traditional painting metaphor with gradients while so zoomed in.}
 
@@
 the zooming and non-zooming cases grow non-linearly over time. However, the non-zooming case takes
-longer for the reasons described. It is not unexpected that even the zooming case grows non-linearly
+longer for the reasons described. \modified{For example, the first ten strokes in both cases take about 2 seconds
+to render. While the last ten strokes take about 20 seconds in the non-zooming case and 7 seconds in the zooming case.}
+It is not unexpected that even the zooming case grows non-linearly
 since even though the geometries involved are not becoming very complicated, the geometry of the entire
 canvas is increasing linearly and must be considered even if the entire canvas is not re-triangulated.
+
+\modified{Note that all these numbers come from running the program on a 1.8 GHz Core i5 MacBook Air with 8GB of RAM and an Intel HD 4000 graphics card.}
 
@@
 this extra geometry the algorithm should achieve a steady growth rate.
+
+\modified{Beyond reducing the complexity of the geometry, the speed of the algorithm might be improved by
+utilizing the GPU. For example, Nehab and Hoppe~\shortcite{Nehab:2008:RRG:1409060.1409088} demonstrate an
+approach for using GPUs to speed up rending of more traditional vector graphics.}
 
@@
 
-\ignorethis{
+\modified{
 \section*{Acknowledgments}
-Omitted for review.
+Let's add something here.
 }

### 2014-06-05 paper.tex
foo

@@
 % uncomment one the following two lines, first one showing pictures
-\documentclass[review]{acmsiggraph}
-%\documentclass[review,draft]{acmsiggraph}
+\documentclass[annual]{acmsiggraph}
+%\documentclass[annual,draft]{acmsiggraph}
 
@@
 
-\author{Mark D.
-Benjamin\thanks{e-mail:mdbenjam@princeton.edu} \hspace{10 pt} Princeton Unviersity\\ Adam Finkelstein\thanks{e-mail:af@cs.princeton.edu} \hspace{10 pt} Princeton University\\ Stephen DiVerdi\thanks{e-mail:diverdi@google.com} \hspace{10 pt} Google}
-\pdfauthor{Mark D. Benjamin, Adam Finkelstein, Stephen Diverdi}
+\author{Mark D. Benjamin \\ Princeton University \and 
+Stephen DiVerdi \\ Google \and 
+Adam Finkelstein \\ Princeton University}
+
+\pdfauthor{Mark D. Benjamin, Stephen DiVerdi, Adam Finkelstein}
 

### 2014-06-06 paper.tex
Reduced the spacing
tweaks to marks' modifications.

@@
 % uncomment one the following two lines, first one showing pictures
-\documentclass[annual]{acmsiggraph}
+\documentclass[conference]{acmsiggraph}
 %\documentclass[annual,draft]{acmsiggraph}
@@
 greedily eliding vertices whenever they lie within a small threshold distance of $P$. \modified{We use
-a threshold relative to zoom of 0.5 which was determined to work well empirically.}
+a threshold of 0.5 pixels at the current zoom level, which works well empirically.}
 
@@
 the limits of single floating point precision; 
-this could be ameliorated by switching to double-precision. \modified{Traditional vector
-graphics software such as Adobe Illustrator caps zooming at a ratio of 64:1. Although this
-is not an inherent limitation, there is currently not a way to use such vector graphics programs
-to draw using a traditional painting metaphor with gradients while so zoomed in.}
+this could be ameliorated by switching to double-precision. \modified{This significantly exceeds the limits
+imposed by existing vector graphics software such as Adobe Illustrator, which has a maximum zoom ratio
+of 64:1.  Furthermore, our ability to make paint strokes of smooth gradients in a flat vector representation (see Figure~\ref{fig:teaser}) is unique.}
 
@@
 longer for the reasons described. \modified{For example, the first ten strokes in both cases take about 2 seconds
-to render. While the last ten strokes take about 20 seconds in the non-zooming case and 7 seconds in the zooming case.}
+to render, while the last ten strokes take about 20 seconds in the non-zooming case and 7 seconds in the zooming case.}
 It is not unexpected that even the zooming case grows non-linearly
@@
 canvas is increasing linearly and must be considered even if the entire canvas is not re-triangulated.
-
-\modified{Note that all these numbers come from running the program on a 1.8 GHz Core i5 MacBook Air with 8GB of RAM and an Intel HD 4000 graphics card.}
 
@@
 \modified{Beyond reducing the complexity of the geometry, the speed of the algorithm might be improved by
-utilizing the GPU. For example, Nehab and Hoppe~\shortcite{Nehab:2008:RRG:1409060.1409088} demonstrate an
-approach for using GPUs to speed up rending of more traditional vector graphics.}
+utilizing the GPU more heavily. For example, Nehab and Hoppe~\shortcite{Nehab:2008:RRG:1409060.1409088}
+demonstrate an approach for using GPUs to speed up rending of more traditional vector graphics.}
 
@@
 \subsection{Performance}
-The current implementation uses Python and PyOpenGL. Python's performance is slower than
+Performance is measured on a 1.8 GHz Core i5 MacBook Air with 8GB of RAM and an Intel
+HD 4000 graphics card.  The current implementation uses Python and PyOpenGL. Python's performance is slower than
 a lower level implementation in C++ would be. The preview rasterization of a new stroke on 
@@
 
-\modified{
-\section*{Acknowledgments}
-Let's add something here.
-}
+%\modified{
+%\section*{Acknowledgments}
+%Let's add something here.
+%}
 

### 2014-06-06 paper.tex
Figure rearrangement.
foo

@@
 Work has also been done in multi-resolution painting using raster graphics. \modified{Early work with Pad++ by Bederson and Hollan~\shortcite{Bederson:1994:PZG:192426.192435} 
-explored the possibility of having an infinite resolution canvas. However, rather than a general painting program 
-this work is focused on creating an interface builder for developers.}
+explored the possibility of having an infinite resolution canvas. However, rather than a painting program, 
+this work focused on creating an interface builder for developers.}
 The method of Berman~\etal~\shortcite{Berman:1994:MPC:192161.192181}
@@
 
+\begin{figure}
+    \centering
+    \setlength{\w}{0.45\columnwidth}
+
+        \includegraphics[width=\w]{images/stroke_triangulation/hardrendered}
+        \includegraphics[width=\w]{images/stroke_triangulation/softrendered}
+
+        \includegraphics[width=\w]{images/stroke_triangulation/hardpruned}
+        \includegraphics[width=\w]{images/stroke_triangulation/softpruned}
+
+        \includegraphics[width=\w]{images/stroke_triangulation/hardmesh}
+        \includegraphics[width=\w]{images/stroke_triangulation/softmesh}
+    \caption{Left: A hard stroke is rasterized, the contour is found, and a triangle mesh is generated.
+             Right: A soft stroke is rasterized in two different shades of gray, two contours are found,
+             a triangle mesh is generated with a boundary separating the inner and outer parts of the stroke.}
+    \label{fig:rastertotriangles}
+\end{figure}
+
 \subsection{Converting Strokes to Contours}
@@
 
+\begin{figure}
+    \centering
+        \fbox{\includegraphics[width=.22\textwidth]{images/emptycanvas}}
+        \fbox{\includegraphics[width=.22\textwidth]{images/canvaswithonestroke}}
+    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
+             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
+             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
+    \label{fig:firststrokes}
+\end{figure}
+
 \subsection{Converting Contours to Triangles}
@@
 approximately represents the drawn stroke.
-
-\begin{figure}
-    \centering
-    \setlength{\w}{0.45\columnwidth}
-
-        \includegraphics[width=\w]{images/stroke_triangulation/hardrendered}
-        \includegraphics[width=\w]{images/stroke_triangulation/softrendered}
-
-        \includegraphics[width=\w]{images/stroke_triangulation/hardpruned}
-        \includegraphics[width=\w]{images/stroke_triangulation/softpruned}
-
-        \includegraphics[width=\w]{images/stroke_triangulation/hardmesh}
-        \includegraphics[width=\w]{images/stroke_triangulation/softmesh}
-    \caption{Left: A hard stroke is rasterized, the contour is found, and a triangle mesh is generated.
-             Right: A soft stroke is rasterized in two different shades of gray, two contours are found,
-             a triangle mesh is generated with a boundary separating the inner and outer parts of the stroke.}
-    \label{fig:rastertotriangles}
-\end{figure}
 
@@
     \centering
-        \fbox{\includegraphics[width=.22\textwidth]{images/emptycanvas}}
-        \fbox{\includegraphics[width=.22\textwidth]{images/canvaswithonestroke}}
-    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
-             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
-             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
-    \label{fig:firststrokes}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
+        \hspace{.2in}
+        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}
+
+    \caption{Left: A red stroke is composited over a blue stroke. An intersection
+            point is inserted in the center and four color arcs must be determined. Right:
+            The four colors are resolved as the four combinations of the red stroke's color
+            arcs composited over the blue stroke's color arcs.}
+    \label{fig:intersectioncolors}
 \end{figure}
+
+\begin{figure*}
+    \centering
+        \includegraphics[width=.32\textwidth]{images/composite21}
+        \includegraphics[width=.32\textwidth]{images/composite22}
+        \includegraphics[width=.32\textwidth]{images/composite23}
+    \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
+             by the the blue stroke and point B is on top of the red stroke. Middle:
+             New color arcs are determined. Both of A's color arcs are composited
+             with the blue stroke which turns red to purple and transparent to blue. Similarly
+             both of B's color arcs are composited on top of the red stroke which turns
+             blue to purple and transparent to red. Two intersection points C and D are
+             added. Their color arcs are determined by examining neighboring points. Right: 
+             A new triangle mesh is generated to preserve boundaries. Note that triangles
+             in the intersection will be purple.}
+    \label{fig:colorarcdeterminer}
+\end{figure*}
 
@@
 the old point must be determined through bilinear interpolation before compositing.
-
-\begin{figure}
-    \centering
-        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
-        \hspace{.2in}
-        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}
-
-    \caption{Left: A red stroke is composited over a blue stroke. An intersection
-            point is inserted in the center and four color arcs must be determined. Right:
-            The four colors are resolved as the four combinations of the red stroke's color
-            arcs composited over the blue stroke's color arcs.}
-    \label{fig:intersectioncolors}
-\end{figure}
-
-\begin{figure*}
-    \centering
-        \includegraphics[width=.32\textwidth]{images/composite21}
-        \includegraphics[width=.32\textwidth]{images/composite22}
-        \includegraphics[width=.32\textwidth]{images/composite23}
-    \caption{Left: A blue stroke is composited over a red stroke. Point A is covered
-             by the the blue stroke and point B is on top of the red stroke. Middle:
-             New color arcs are determined. Both of A's color arcs are composited
-             with the blue stroke which turns red to purple and transparent to blue. Similarly
-             both of B's color arcs are composited on top of the red stroke which turns
-             blue to purple and transparent to red. Two intersection points C and D are
-             added. Their color arcs are determined by examining neighboring points. Right: 
-             A new triangle mesh is generated to preserve boundaries. Note that triangles
-             in the intersection will be purple.}
-    \label{fig:colorarcdeterminer}
-\end{figure*}
 
@@
 the limits of single floating point precision; 
-this could be ameliorated by switching to double-precision. \modified{This significantly exceeds the limits
+this could be ameliorated by switching to double-precision. 
+\modified{Nevertheless, this range far exceeds limits
 imposed by existing vector graphics software such as Adobe Illustrator, which has a maximum zoom ratio
-of 64:1.  Furthermore, our ability to make paint strokes of smooth gradients in a flat vector representation (see Figure~\ref{fig:teaser}) is unique.}
+of 64:1.  Furthermore, our ability to make paint strokes of smooth gradients in a flat vector representation (\eg, Figure~\ref{fig:teaser}) is unique.}
 

### 2014-06-06 paper.tex
Final edit, per Adam's suggestion.

@@
 imposed by existing vector graphics software such as Adobe Illustrator, which has a maximum zoom ratio
-of 64:1.  Furthermore, our ability to make paint strokes of smooth gradients in a flat vector representation (\eg, Figure~\ref{fig:teaser}) is unique.}
+of 64:1.  Furthermore, the ability to make paint strokes of smooth gradients in a flat vector representation (\eg, Figure~\ref{fig:teaser}) differentiates our approach from previous methods.}
 

### 2014-06-06 paper.tex
Answered the two questions and changed render to process

@@
 First, the user's mouse movements are captured by a set of polygons which can be
-rendered to yield a rasterized version of the stroke. \modified{This rasterization happens at the resolution of the current window. Therefore
+rendered to yield a rasterized version of the stroke. \modified{This rasterization happens at the resolution of the current window, 1280x960. Therefore
 whatever the artist sees on the screen is the basis for the vectorization.}
@@
 must cover the same area of canvas over and over. This
-results in more geometry and therefore takes longer to render. The zooming case explores different
+results in more geometry and therefore takes longer to \modified{process this new stroke}. The zooming case explores different
 areas of the canvas and therefore results in less overlapping geometry. This situation should result
@@
 
-Figure~\ref{fig:timing} gives a similar result for the cumulative time strokes take to render. Both
+Figure~\ref{fig:timing} gives a similar result for the cumulative time strokes take to process \modified{after mouse up}. Both
 the zooming and non-zooming cases grow non-linearly over time. However, the non-zooming case takes
-longer for the reasons described. \modified{For example, the first ten strokes in both cases take about 2 seconds
-to render, while the last ten strokes take about 20 seconds in the non-zooming case and 7 seconds in the zooming case.}
+longer for the reasons described. \modified{For example, the first ten strokes in both cases take about 2 seconds each
+to process, while the last ten strokes take about 20 seconds each in the non-zooming case and 7 seconds each in the zooming case.}
 It is not unexpected that even the zooming case grows non-linearly
@@
 one second to process. However, on a canvas with lots of geometry a stroke that overlaps
-that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
+that geometry can take several seconds to process. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
 necessary for the old points affected by the new stroke and the points associated with the
@@
     \includegraphics[height=1.2in]{graphs/cumulativetime}
-    \caption{Rendering time}
+    \caption{Processing time}
     \label{fig:timing}
   \end{subfigure}%
-  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting.
+  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative process time of the painting.
 The blue points are the non-zooming case, while the red points are the zooming case.}

### 2014-06-06 paper.tex
Changed modified to be not green. Changed the word render to process in the graph and text.

@@
 must cover the same area of canvas over and over. This
-results in more geometry and therefore takes longer to \modified{process this new stroke}. The zooming case explores different
+results in more geometry and therefore takes longer to \modified{process}. The zooming case explores different
 areas of the canvas and therefore results in less overlapping geometry. This situation should result
@@
   \end{subfigure}%
-  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative process time of the painting.
+  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative processing time of the painting.
 The blue points are the non-zooming case, while the red points are the zooming case.}
